---
title: "ARO_Response_1"
output: pdf_document
date: "2026-01-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/home/engerlab-2/Documents/gene_set_study/")
```

#setup
```{r}
library(tidyverse)
library(survival)
library(rms)
```

if doing gene set analysis, run this chunk first, then run the gene section, then unload these libraries (they interfere with tidyverse)
```{r}
library(singscore)
library(GSEABase)
library(msigdb)
library(ExperimentHub)
```

#process mRNA data
##load hallmark genes
```{r}
eh = ExperimentHub()
query(eh , 'msigdb')
msigdb.hs = getMsigdb(org = 'hs', id = 'SYM', version = '7.4')
msigdb.hs
listCollections(msigdb.hs)
hallmarks <- subsetCollection(msigdb.hs, 'h')
```

check redundancy
```{r}
genes <- c()
for (i in 1:50)
{
  hallmark <- hallmarks[i][[1]]@geneIds  #check the actual sets we want 
  genes <- c(genes, hallmark)
}
x <- sort(table(genes), decreasing = T)
x['TP53']
table(table(genes))
```

##apply singscore algorithm in batches
(this requires your mRNA files downloaded locally)
```{r warning = FALSE}
fraction_dataframe <- data.frame()
setwd("./m11_zip")
folders <- list.files()
length(folders)
index = 1
for (folder in folders)
{
  print(index)
  setwd(folder)
  patient_id <- folder
  tsv_file <- list.files(pattern = "\\.tsv$", full.names = TRUE, recursive = TRUE)[1]
  test_tsv <- read.table(tsv_file, sep = '\t', header = TRUE)
  
  length_tsv <- length(test_tsv$gene_name)
  test_tsv_mod <- test_tsv %>% dplyr::select(gene_name, tpm_unstranded)
  test_tsv_mod <- test_tsv_mod[5:length_tsv, ]
  test_tsv_mod <- test_tsv_mod %>% distinct(gene_name, .keep_all = TRUE)
  
  test_tsv_1col <- data.frame(test_tsv_mod$tpm_unstranded)
  rownames(test_tsv_1col) <- test_tsv_mod$gene_name
  test_tsv_1col_filter <- test_tsv_1col %>% filter(test_tsv_mod.tpm_unstranded != 0)
  rankData_test <- rankGenes(test_tsv_1col_filter)
  
  names <- c()
  scores_dir <-  c()
  scores_undir <-  c()
  dispersions_dir <- c()
  dispersions_undir <- c()
  
  for (hallmarkNumber in 1:50)
  {
    h1 <- hallmarks[hallmarkNumber] #get the sets we want
    h1data <- h1[[1]] #this is a gene set
    name <- GSEABase::setName(h1data)
    score_dir <- simpleScore(rankData_test, h1data, knownDirection = TRUE)$TotalScore
    score_undir <- simpleScore(rankData_test, h1data, knownDirection = FALSE)$TotalScore
    dispersion_dir <- simpleScore(rankData_test, h1data, knownDirection = TRUE)$TotalDispersion
    dispersion_undir <- simpleScore(rankData_test, h1data, knownDirection = FALSE)$TotalDispersion
    names <- c(names, name)
    scores_dir <- c(scores_dir, score_dir)
    scores_undir <- c(scores_undir, score_undir)
    dispersions_dir <- c(dispersions_dir, dispersion_dir)
    dispersions_undir <- c(dispersions_undir, dispersion_undir)
  }
  

  #add to combined dataframe 
  data_list <- c(patient_id)
  names(data_list) <- "Patient_ID"
  for (i in 1:50) {
    data_list[paste0(names[i], "_Score_Dir")] <- scores_dir[i]
    data_list[paste0(names[i], "_Score_Undir")] <- scores_undir[i]
    data_list[paste0(names[i], "_Dispersion_Dir")] <- dispersions_dir[i]
    data_list[paste0(names[i], "_Dispersion_Undir")] <- dispersions_undir[i]
  }
  
  df <- as.data.frame(t(data_list), stringsAsFactors = FALSE)
  df[-1] <- lapply(df[-1], as.numeric)
  fraction_dataframe <- rbind(fraction_dataframe, df)
  
  setwd("..")
  
  index = index + 1
}

write.csv(fraction_dataframe, "m11_zip_extracted.csv")
```

##merge batches
```{r}
big_df <- data.frame()
setwd("./outputs")
files = list.files()
for (file in files)
{
  temp_csv <- read.csv(file)
  big_df <- rbind(big_df, temp_csv)
}

write.csv(big_df, "full_gene_data.csv")
```

##add patient IDs
```{r}
file_mappings <- read.csv("json_file_mappings.csv")
file_mappings <- file_mappings %>% rename(case_id = patients)
file_mappings <- file_mappings %>% mutate(case_id = substr(case_id, start = 1, stop = 12))

big_df <- read.csv("full_gene_data.csv")
big_df <- big_df[, -c(1,2)]
big_df <- big_df %>% rename(files = Patient_ID)

json_merged <- left_join(big_df, file_mappings, by = "file_ids")

write.csv(json_merged, "gene_data_with_ids.csv")
```


##replace above code for RSI calculation

Gene expressions for the 10 genes rank-ordered, with lowest expression ranked 1.

then apply: 
RSI = 
- 0.0098009 * AR 
+ 0.0128283 * JUN
+ 0.0254552 * STAT1
- 0.0017589 * PRKCB 
- 0.0038171 * RELA
+ 0.1070213 * ABL1
- 0.0002509 * SUMO1
- 0.0092431 * CDK1 
- 0.0204469 * HDAC1 
- 0.0441683 * IRF1

```{r warning = FALSE}
fraction_dataframe <- data.frame()
setwd("./m12_zip")
folders <- list.files()
length(folders)
index = 1
for (folder in folders)
{
  if(index %% 100 == 0){print(index)}
  setwd(folder)
  patient_id <- folder
  tsv_file <- list.files(pattern = "\\.tsv$", full.names = TRUE, recursive = TRUE)[1]
  test_tsv <- read.table(tsv_file, sep = '\t', header = TRUE)
  
  length_tsv <- length(test_tsv$gene_name)
  test_tsv_mod <- test_tsv %>% dplyr::select(gene_name, tpm_unstranded)
  test_tsv_mod <- test_tsv_mod[5:length_tsv, ]
  test_tsv_mod <- test_tsv_mod %>% distinct(gene_name, .keep_all = TRUE)
  
  test_tsv_1col <- data.frame(test_tsv_mod$tpm_unstranded)
  rownames(test_tsv_1col) <- test_tsv_mod$gene_name
  test_tsv_1col_filter <- test_tsv_1col %>% filter(test_tsv_mod.tpm_unstranded != 0)
  
  
  rsi_genes <- c("AR", "JUN", "STAT1", "PRKCB", "RELA", "ABL1", "SUMO1", "CDK1", "HDAC1", "IRF1")
  rsi_coefs <- c(-0.0098009, 0.0128283, 0.0254552, -0.0017589, -0.0038171, 0.1070213, -0.0002509, -0.0092431, -0.0204469, 0.0441683)
  seq_10 <- seq(1,10)
  exprs <- c()
  for(gene in rsi_genes)
  {
    expr <- test_tsv_1col[gene,]
    exprs <- c(exprs, expr)
  }
  
  comb <- data.frame(rsi_genes, rsi_coefs, exprs)
  comb <- comb %>% arrange(exprs)
  comb <- cbind(comb, seq_10)
  comb <- comb %>% mutate(val = seq_10 * rsi_coefs)
  rsi <- sum(comb$val)
  df <- data.frame("Patient_ID" = patient_id, "RSI" = rsi)
  
  fraction_dataframe <- rbind(fraction_dataframe, df)
  
  setwd("..")
  
  index = index + 1
}
```

#create data files for further processing
##clinical variables
```{r}
mmc_data <- read.csv("mmc1.csv")
mmc_small <- mmc_data %>% select(bcr_patient_barcode,type, age_at_initial_pathologic_diagnosis, gender, ajcc_pathologic_tumor_stage, clinical_stage, histological_type, histological_grade, initial_pathologic_dx_year, PFI, PFI.time, OS, OS.time)
#for LAML, no PFI, so replace with OS
mmc_small <- mmc_small %>% mutate(PFI = ifelse(type == "LAML", OS, PFI))
mmc_small <- mmc_small %>% select(-c(OS, OS.time))
mmc_mut <- mmc_small
mmc_mut <- mmc_mut %>% rename(age = age_at_initial_pathologic_diagnosis, path_stage = ajcc_pathologic_tumor_stage, year = initial_pathologic_dx_year, case = bcr_patient_barcode)
```

##missingness
parse and merge missingness indicators
```{r}
mmc_mut <- mmc_mut %>% mutate(
    type = ifelse(type == "#N/A", NA, type), 
    age = ifelse(age == "#N/A", NA, age), 
    gender = ifelse(gender == "#N/A", NA, gender), 
    path_stage = ifelse(path_stage == "#N/A", NA, path_stage), 
    clinical_stage = ifelse(clinical_stage == "#N/A", NA, clinical_stage), 
    histological_type = ifelse(histological_type == "#N/A", NA, histological_type), 
    histological_grade = ifelse(histological_grade == "#N/A", NA, histological_grade), 
    year = ifelse(year == "#N/A", NA, year), 
    PFI = ifelse(PFI == "#N/A", NA, PFI), 
    PFI.time = ifelse(PFI.time == "#N/A", NA, PFI.time)
)
mmc_mut <- mmc_mut %>% mutate(
  path_stage = ifelse(path_stage %in% c("[Not Available]", "[Not Applicable]", "[Discrepancy]", "[Unknown]", "Stage X"), NA, path_stage), 
  clinical_stage = ifelse(clinical_stage %in% c("[Not Available]", "[Not Applicable]", "[Discrepancy]", "[Unknown]"), NA, clinical_stage)
)
```

##create 4 stage categories
```{r}

stage1path <- c("I/II NOS", "IS", "Stage I", "Stage IA", "Stage IB")
stage2path <- c("Stage II", "Stage IIA", "Stage IIB", "Stage IIC")
stage3path <- c("Stage III", "Stage IIIA", "Stage IIIB", "Stage IIIC")
stage4path <- c("Stage IV", "Stage IVA", "Stage IVB", "Stage IVC")

stage1clin <- c("I", "Stage I", "Stage IA", "Stage IA1", "Stage IA2", "Stage IB", "Stage IB1", "Stage IB2", "Stage IC", "Stage IS")
stage2clin <- c("IIa", "IIb", "Stage II", "Stage IIA", "Stage IIA1", "Stage IIA2", "Stage IIB", "Stage IIC")
stage3clin <- c("III", "Stage III", "Stage IIIA", "Stage IIIB", "Stage IIIC", "Stage IIIC1", "Stage IIIC2")
stage4clin <- c("IVa", "IVb", "Stage IV", "Stage IVA", "Stage IVB", "Stage IVC")


mmc_mut <- mmc_mut %>% mutate( 
  path_stage = ifelse(path_stage %in% stage1path, "Stage 1", 
                      ifelse(path_stage %in% stage2path, "Stage 2", 
                             ifelse(path_stage %in% stage3path, "Stage 3", 
                                    ifelse(path_stage %in% stage4path, "Stage 4", NA)))))
mmc_mut <- mmc_mut %>% mutate( 
  clinical_stage = ifelse(clinical_stage %in% stage1clin, "Stage 1", 
                      ifelse(clinical_stage %in% stage2clin, "Stage 2", 
                             ifelse(clinical_stage %in% stage3clin, "Stage 3", 
                                    ifelse(clinical_stage %in% stage4clin, "Stage 4", NA)))))
mmc_mut <- mmc_mut %>% mutate(merged_stage = ifelse(is.na(path_stage), clinical_stage, path_stage))                        
```

```{r}
mmc_premice <- mmc_mut
mmc_premice_save <- write.csv(mmc_premice, "mmc_premice_save.csv")
```
gene_data_all
##select primary tumours
```{r}
gene_data_ids <- read.csv("gene_data_with_ids.csv") #made above in first steps
manifest_full <- read_tsv("tumor_type_manifest.txt") #manifest of primary tumours
gene_ids_tumour <- gene_data_ids %>% filter(file_ids %in% manifest_full$id)
gene_ids_tumour <- gene_ids_tumour %>% select(-c(X, file_ids, ids))
```

##clinical treatments
```{r}
full_clinical_data <- read.csv("full_clinical_file.csv")
full_clinical_data <- full_clinical_data %>% rename(case = case_submitter_id)

full_clinical_radio <- full_clinical_data %>% filter(treatment_type == "Radiation Therapy, NOS")
full_clinical_radio <- full_clinical_radio %>% select(case, treatment_or_therapy)
full_clinical_radio <- full_clinical_radio %>% rename(case_id = case, radiotherapy = treatment_or_therapy)

full_clinical_chemo <- full_clinical_data %>% filter(treatment_type == "Pharmaceutical Therapy, NOS")
full_clinical_chemo <- full_clinical_chemo %>% select(case, treatment_or_therapy)
full_clinical_chemo <- full_clinical_chemo %>% rename(case_id = case, chemotherapy = treatment_or_therapy)

full_clinical_merged <- left_join(full_clinical_radio, full_clinical_chemo, by = join_by(case_id))

gene_data_merged <- left_join(full_clinical_merged, gene_ids_tumour, by = join_by(case_id))
```

##process
merge gene data across multiple samples
###genes
```{r}
gene_data_all <- gene_data_merged %>% rename(case = case_id)
gene_data_all <- gene_data_all %>% group_by(case) %>% 
  summarise(
    radiotherapy = first(radiotherapy),
    chemotherapy = first(chemotherapy),
    across(where(is.numeric), mean, na.rm = TRUE)) #am averaging all TUMOUR samples here

gene_data_all <- gene_data_all %>% mutate(radiotherapy = ifelse(radiotherapy == "not reported", NA, radiotherapy), chemotherapy = ifelse(chemotherapy == "not reported", NA, chemotherapy))
```

###missingness
```{r}
mmc_premice <- read.csv("mmc_premice_save.csv")
mmc_premice <- mmc_premice %>% select(-c(X))

pre_mice_dataframe <- left_join(gene_data_all, mmc_premice, by = join_by(case))

#exclude missing outcomes
pre_mice_dataframe <- pre_mice_dataframe %>% filter(!is.na(PFI)) %>% filter(!is.na(PFI.time)) 

#exclude missing genes
pre_mice_dataframe <- pre_mice_dataframe %>% filter(!is.na(HALLMARK_TNFA_SIGNALING_VIA_NFKB_Score_Dir))

pre_mice_rt <- pre_mice_dataframe %>% filter(radiotherapy == "yes")
pre_mice_rt_complete <- pre_mice_rt %>% drop_na(any_of(c("type", "merged_stage", "age", "gender")))
```

11080 patients with clinical data
10907 after deleting missing outcomes = 173 no outcomes
10271 with genomic data = 636 no gene expression
3269 with confirmed radiotherapy = 7002 no RT or 
2797 after missingness for age, stage, or gender = 472 incomplete data

###grade (for lgg)
```{r}
#process grade
hist_na <- c("[Discrepancy]", "[Not Available]", "[Unknown]", "GX", "GB")
table(pre_mice_dataframe$histological_grade)
pre_mice_dataframe <- pre_mice_dataframe %>% mutate(histological_grade = ifelse(histological_grade %in% hist_na, NA, ifelse(histological_grade == "Low Grade", "G1", ifelse(histological_grade == "High Grade", "G3", histological_grade))))
#Low and High refer only to BLCA - Low grade is the same as grade 1. High grade is the same as grade 3.

#change types
pre_mice_dataframe <- pre_mice_dataframe %>% mutate(type = as.factor(type), age = as.numeric(age), gender = as.factor(gender), path_stage = as.factor(path_stage), clinical_stage = as.factor(clinical_stage), histological_grade = as.factor(histological_grade), year = as.numeric(year), PFI = as.numeric(PFI), PFI.time = as.numeric(PFI.time), merged_stage = as.factor(merged_stage))

#to include LGG, use grade instead of stage
pre_mice_complete <- pre_mice_dataframe %>% mutate(histological_grade = ifelse(histological_grade == "G1", "Stage 1", ifelse(histological_grade == "G2", "Stage 2", ifelse(histological_grade == "G3", "Stage 3", ifelse(histological_grade == "G4", "Stage 4", NA)))))
pre_mice_complete <- pre_mice_complete %>% mutate(merged_stage = as.character(merged_stage))
pre_mice_complete <- pre_mice_complete %>% mutate(merged_stage = ifelse(type == "LGG", histological_grade, merged_stage))
pre_mice_complete <- pre_mice_complete %>% mutate(merged_stage = as.factor(merged_stage))

#clean 
pre_mice_complete <- pre_mice_complete %>% select(-c(clinical_stage, path_stage))
pre_mice_complete <- pre_mice_complete %>% select(!contains("Undir"))
```


##fix data types
```{r}
write.csv(pre_mice_complete, "pre_mice_complete_oct_20.csv")
pre_mice <- read.csv("pre_mice_complete_oct_20.csv")
pre_mice <- pre_mice %>% select(-c(X, X.1))
pre_mice <- pre_mice %>% mutate(type = as.factor(type), age = as.numeric(age), gender = as.factor(gender), histological_grade = as.factor(histological_grade), year = as.numeric(year), PFI = as.numeric(PFI), PFI.time = as.numeric(PFI.time), merged_stage = as.factor(merged_stage), radiotherapy = as.factor(radiotherapy), chemotherapy = as.factor(chemotherapy))
```

#patient characteristics
##table 1 - patient characteristics

```{r}
pre_mice_rt <- pre_mice %>% filter(radiotherapy == "yes")
pre_mice_rt_complete <- pre_mice_rt %>% drop_na(any_of(c("type", "merged_stage", "age", "gender")))
sort(table(pre_mice_rt$type))
```


```{r}
examine_types <-  c("BRCA", "CESC", "HNSC", "LGG", "LUAD","SKCM", "THCA", "UCEC")
tbl1_df <- pre_mice_rt %>% mutate(type = as.character(type))
tbl1_df <- tbl1_df %>% mutate(type = ifelse(type %in% examine_types, type, "OTHER"))
tbl1_df <- tbl1_df %>% mutate(PFI.time != 0)
```


```{r}
to_make_tbl1 <- c("BRCA", "CESC", "HNSC", "LGG", "LUAD","SKCM", "THCA", "UCEC", "OTHER")
df <- data.frame()
for (typeC in to_make_tbl1)
{
  typeData <- tbl1_df %>% filter(type == typeC)
  complete <- typeData %>% drop_na(any_of(c("age", "merged_stage", "gender")))
  complete_n <- nrow(complete)
  # print(typeC)
  # print(table(typeData$PFI))
  # print(mean(typeData$PFI.time))
  # print(sd(typeData$PFI.time))
  # print(mean(typeData$age))
  # print(sd(typeData$age))
  # print(table(typeData$merged_stage))
  # print(table(typeData$gender))
  samp_low = round(sum(typeData$PFI)/20)
  samp_high = round(sum(typeData$PFI)/10)
  df_type <- data.frame(Total_N = nrow(typeData), 
                        PFI_1 =  table(typeData$PFI)['1'], 
                        PFI_0 = table(typeData$PFI)['0'], 
                        PFI_mean_sd = paste(round(mean(typeData$PFI.time), 2), " (", round(sd(typeData$PFI.time), 2), ")", sep = ""),
                        #samp_size = paste(deg_free, " (", epp, ")", sep = ""), 
                        samp_size = paste(samp_low, " - ", samp_high, sep = ""),
                        complete_size = complete_n,
                        age_mean_sd = paste(round(mean(typeData$age, na.rm = T), 2), " (", round(sd(typeData$age, na.rm = T), 2), ")", sep = ""), 
                        stage_1 = table(typeData$merged_stage)['Stage 1'], 
                        stage_2 = table(typeData$merged_stage)['Stage 2'], 
                        stage_3 = table(typeData$merged_stage)['Stage 3'], 
                        stage_4 = table(typeData$merged_stage)['Stage 4'], 
                        gender_M =  table(typeData$gender)['MALE'], 
                        gender_F = table(typeData$gender)['FEMALE'])
  df <- rbind(df, df_type)
}

  complete <- tbl1_df %>% drop_na(any_of(c("age", "merged_stage", "gender")))
  complete_n <- nrow(complete)

  samp_low = round(sum(tbl1_df$PFI)/20)
  samp_high = round(sum(tbl1_df$PFI)/10)
  df_all <- data.frame(Total_N = nrow(tbl1_df), 
                       PFI_1 =  table(tbl1_df$PFI)['1'], 
                        PFI_0 = table(tbl1_df$PFI)['0'], 
                        PFI_mean_sd = paste(round(mean(tbl1_df$PFI.time), 2), " (", round(sd(tbl1_df$PFI.time), 2), ")", sep = ""),
                        samp_size = paste(samp_low, " - ", samp_high, sep = ""),
                        complete_size = complete_n,
                        age_mean_sd = paste(round(mean(tbl1_df$age, na.rm = T), 2), " (", round(sd(tbl1_df$age, na.rm = T), 2), ")", sep = ""), 
                        stage_1 = table(tbl1_df$merged_stage)['Stage 1'], 
                        stage_2 = table(tbl1_df$merged_stage)['Stage 2'], 
                        stage_3 = table(tbl1_df$merged_stage)['Stage 3'], 
                        stage_4 = table(tbl1_df$merged_stage)['Stage 4'], 
                        gender_M =  table(tbl1_df$gender)['MALE'], 
                        gender_F = table(tbl1_df$gender)['FEMALE'])
  df <- rbind(df, df_all)
  
rownames_tbl1 <- c("BRCA", "CESC", "HNSC", "LGG", "LUAD","SKCM", "THCA", "UCEC", "OTHER", "ALL")
rownames(df) = rownames_tbl1
df
```

```{r}
getwd()
write.csv(df, "table1.csv")
```

##figure 1 - genes heat map

```{r}
pre_mice_rt <- pre_mice %>% filter(radiotherapy == "yes")
pre_mice_rt_complete <- pre_mice_rt %>% drop_na(any_of(c("type", "merged_stage", "age", "gender")))
all_genes <- colnames(pre_mice)[4:103]
```


```{r}
short_all_genes <- c()
for (i in 1:100)
{
  str = substr(all_genes[i], 10, nchar(all_genes[i])-4)
  str <- gsub("_", " ", str)
  short_all_genes <- c(short_all_genes, str)
}

gene_map <- setNames(short_all_genes, all_genes)
```


```{r}
examine_types <-  c("BRCA", "CESC", "HNSC", "LGG", "LUAD","SKCM", "THCA", "UCEC")
fig_1_df <- pre_mice_rt %>% mutate(type = as.character(type))
fig_1_df <- fig_1_df %>% mutate(PFI.time != 0)
fig_1_df <- fig_1_df %>% mutate(type = ifelse(type %in% examine_types, type, "OTHER"))

#scale across all genes
df_scaled <- fig_1_df %>%
  mutate(across(
    all_of(all_genes),
    ~ scale(.)[,1])) 

#summarize by type and PFI
gene_means <- df_scaled %>%
  group_by(type, PFI) %>%
  summarise(across(all_of(all_genes), mean, na.rm = TRUE), .groups = "drop")

#summarize just by outcome (for final column)
all_means <- df_scaled %>%
  group_by(PFI) %>%
  summarise(across(all_of(all_genes), mean, na.rm = TRUE), .groups = "drop") %>%
  mutate(type = "ALL")

gene_means <- bind_rows( #combine those two
  gene_means,   
  all_means
)

gene_means_long <- gene_means %>%
  pivot_longer(cols = all_of(all_genes),
               names_to = "gene",
               values_to = "mean_expr")

gene_means_long <- gene_means_long %>% #make good labels
    mutate(
    type_outcome = paste(type, PFI, sep = ": "),
    type_outcome = factor(
      type_outcome,
      levels = expand.grid(
        type = sort(unique(type)),
        PFI  = sort(unique(PFI))
      ) %>%
        transmute(label = paste(type, PFI, sep = ": ")) %>%
        pull(label)
    )
  )

#("BRCA", "CESC", "HNSC", "LGG", "LUAD","SKCM", "THCA", "UCEC")
gene_means_long$type_outcome <- factor( #set order
  gene_means_long$type_outcome,
  levels = c(
    "BRCA: 0", "BRCA: 1",
    "CESC: 0", "CESC: 1",
    "HNSC: 0", "HNSC: 1", 
     "LGG: 0", "LGG: 1", 
     "LUAD: 0", "LUAD: 1", 
     "SKCM: 0", "SKCM: 1", 
     "THCA: 0", "THCA: 1", 
    "UCEC: 0", "UCEC: 1",
     "OTHER: 0", "OTHER: 1", 
    "ALL: 0", "ALL: 1"
    
  )
)

gene_means_long <- gene_means_long %>% #gene labels
  mutate(gene = gene_map[gene])

gene_means_long <- gene_means_long %>% 
  mutate(gene = factor(gene, levels = sort(unique(gene), decreasing = T)))

line_positions <- seq(0.5, by = 2, length.out = 9 + 2) #make black lines

gene_order <- gene_means_long %>% #sort genes
  filter(type == "ALL") %>%
  dplyr::select(gene, PFI, mean_expr) %>%
  pivot_wider(names_from = PFI, values_from = mean_expr) %>%
  mutate(diff = `1` - `0`) %>%
  arrange((diff)) %>%
  pull(gene)

gene_means_long$gene <- factor(
  gene_means_long$gene,
  levels = gene_order
)

ggplot(gene_means_long, aes(x = type_outcome, y = gene, fill = mean_expr)) + #make the plot
  geom_tile() +
  geom_vline(
    xintercept = line_positions,
    color = "black",
    linewidth = 0.6
  ) +
  scale_fill_viridis_c(option = "B") +  # perceptually uniform
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.text.y = element_text(size = 6)  # reduce if 100 genes
  ) +
  labs(
    x = "Cancer Type & Outcome",
    y = "Gene",
    fill = "Standardized Mean Value",
    title = "Standardized Mean Gene Set Parameter Value by Cancer Type and Outcome"
  )

ggsave("gene_plot.pdf", width = 12, height = 16)
ggsave("gene_plot.png", width = 12, height = 16)
```


#analysis
##full pan-cancer fit

```{r}
pre_mice_rt <- pre_mice %>% filter(radiotherapy == "yes")
complete_cases_rt <- pre_mice_rt
complete_cases_rt <- complete_cases_rt %>% filter(PFI.time != 0)
complete_cases_rt <- complete_cases_rt %>% mutate(type = as.character(type))
complete_cases_rt <- complete_cases_rt %>% mutate(type = ifelse(type %in% examine_types, type, "OTHER"))
complete_cases_rt <- complete_cases_rt %>% mutate(type = as.factor(type))


scaled_df <- complete_cases_rt %>% mutate(across(all_genes, ~scale(., center = T, scale = T)))
formula_all <- as.formula(paste("Surv(PFI.time, PFI) ~", paste(all_genes,  collapse = " + "), "+ type"))
cox_all <- coxph(formula_all, data = scaled_df)
summary(cox_all)

```

```{r}
cox.zph(cox_all)
plot(cox.zph(cox_all))
plot(predict(cox_all), residuals(cox_all, type="martingale"), 
     xlab="Predicted Value", ylab="Martingale Residuals")
ggcoxdiagnostics(cox_all, type = , linear.predictions = TRUE)
ggcoxdiagnostics(cox_all, type = "dfbeta",
                 linear.predictions = FALSE, ggtheme = theme_bw())
```


```{r}
#plot pan cancer coxph
beta <- coef(cox_all)
ci   <- confint(cox_all)
pval <- summary(cox_all)$coefficients[, "Pr(>|z|)"]

names <- names(beta)
names_short <- c()
for (i in 1:length(names))
{
  if(names[i] %in% all_genes)
    {
      str = substr(names[i], 10, nchar(names[i])-4)
      str <- gsub("_", " ", str)
      names_short <- c(names_short, str)
    }
  else
  {
    names_short <- c(names_short, names[i] )
  }
}

# Build final table (HR scale, rounded)
final_table <- data.frame(
  variable = names_short,
  estimate = round(exp(beta), 2),
  CI_low   = round(exp(ci[, 1]), 2),
  CI_high  = round(exp(ci[, 2]), 2),
  pval = pval,
 # p_value = formatC(pval, format = "e", digits = 2),
  row.names = NULL
)

final_table$p_value_fmt <- ifelse(final_table$pval < 0.001, "p < 0.001",
                         paste("p = ", formatC(final_table$pval, format = "f", digits = 3), sep = ""))
# final_table[] <- lapply(final_table, function(x)
#   if (is.numeric(x)) formatC(x, format = "f", digits = 2) else x
# )

final_table <- final_table %>% mutate(toPrint = paste(estimate, " [", CI_low, " - ", CI_high, "] ", p_value_fmt, sep = ""))
final_table <- final_table %>% select(variable, toPrint)

# Write to CSV
write.csv(final_table, "pan_cancer_coxph_results.csv", row.names = FALSE)

```


##KM

```{r}
library(ggsurvfit)
library(survminer)
library(caret)

examine_types <-  c("BRCA", "CESC", "HNSC", "LGG", "LUAD","SKCM", "THCA", "UCEC")

pre_mice_rt <- pre_mice %>% filter(radiotherapy == "yes")
complete_cases_rt <- pre_mice_rt #%>% drop_na(any_of(c("type", "merged_stage", "age", "gender")))
complete_cases_rt <- complete_cases_rt %>% filter(PFI.time != 0)

complete_cases_rt <- complete_cases_rt %>% mutate(type = as.character(type))
complete_cases_rt <- complete_cases_rt %>% mutate(type = ifelse(type %in% examine_types, type, "OTHER"))
complete_cases_rt <- complete_cases_rt %>% mutate(type = as.factor(type))

merged_predictions <- data.frame()
plots = list()

i = 1
for (typeC in examine_types)
{
  type_data <- complete_cases_rt %>% filter(type == typeC) #cases of the evaluation type
  non_type_data <- complete_cases_rt %>% filter(type != typeC) #all other cases
  
  formula_all <- as.formula(paste("Surv(PFI.time, PFI) ~", paste(all_genes,  collapse = " + "), "+ type"))
  cox_pan_cancer <- coxph(formula_all, data = non_type_data)
    
  test_pan_predictions <- predict(cox_pan_cancer, type = "lp", newdata = type_data)
    test_pan_predictions <- as.numeric(test_pan_predictions)
    type_data$pan_predictions <- test_pan_predictions
  

  plot_predictions <- type_data
  median_pan <- median(plot_predictions$pan_predictions)
  plot_predictions <- plot_predictions %>% mutate(binary_pan = ifelse(pan_predictions < median_pan, 0, 1))
  plot_predictions <- plot_predictions %>% mutate(PFI.time = PFI.time / 365.25) #put time into years
  fit <- survfit(Surv(PFI.time, PFI) ~ binary_pan, data = plot_predictions)
  survPlot <- (ggsurvplot(fit, conf.int = T, pval = T, risk.table = T, tables.height = 0.3, xlim = c(0,5), break.time.by = 0.5, xlab = "Time (years)", legend.title = "Pan-Cancer Risk", legend.labs = c("Low Risk", "High Risk"), title = typeC)) #legend = none so we can make shared legend later

  plots[[i]] = survPlot
 print(survPlot) 
 
   merged_predictions <- rbind(merged_predictions, type_data)
   
  individual_cox <- coxph(Surv(PFI.time, PFI) ~ pan_predictions, data = type_data)
  print(typeC)
  print(summary(individual_cox))
  
       type_data <- type_data %>% drop_na(any_of(c("type", "merged_stage", "age", "gender"))) 
   
      if (typeC == "BRCA" || typeC == "CESC" || typeC == "UCEC") #sex-specific cancers, no gender var
     {
          individual_cox_2 <- coxph(Surv(PFI.time, PFI) ~ age + merged_stage + pan_predictions, data = type_data)
      }else
      {
        individual_cox_2 <- coxph(Surv(PFI.time, PFI) ~ age + gender + merged_stage + pan_predictions, data = type_data)
      }

  i = i + 1
}
```


```{r}
surv_list <- plots
plot_list <- lapply(surv_list, function(x) x$plot)

combined_plot <- arrange_ggsurvplots(
  surv_list,
  ncol = 2,
  nrow = 4,
  risk.table.height = 0.25,
  legend = "bottom"
)

ggsave(
  "survival_panels_with_tables.pdf",
  combined_plot,
  device = cairo_pdf,
  width = 12,
  height = 24,
  units = "in"
)

ggsave(
  "survival_panels_with_tables.png",
  combined_plot,
  device = "png",
  width = 12,
  height = 24,
  units = "in"
)
```


###merged predictions boxplots
```{r}
merged_predictions %>%
  mutate(
    type = factor(type),
    PFI  = factor(PFI, levels = c(0, 1), labels = c("PFI = 0", "PFI = 1"))
  ) %>%
  ggplot(aes(x = type, y = pan_predictions, fill = PFI)) +
  geom_boxplot(
    position = position_dodge(width = 0.8),
    outlier.alpha = 0.4
  ) +
  labs(
    x = "Cancer type",
    y = "Linear Predictions",
    fill = "Outcome"
  ) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.x = element_blank()
  )

```

"The lower and upper hinges correspond to the first and third quartiles (the 25th and 75th percentiles)."
"The upper whisker extends from the hinge to the largest value no further than 1.5 * IQR from the hinge (where IQR is the inter-quartile range, or distance between the first and third quartiles). The lower whisker extends from the hinge to the smallest value at most 1.5 * IQR of the hinge. Data beyond the end of the whiskers are called "outlying" points and are plotted individually."

####with wixocon test
```{r}
library(ggpubr)
merged_predictions %>%
  mutate(
    type = factor(type),
    PFI  = factor(PFI, levels = c(0, 1), labels = c("PFI = 0", "PFI = 1"))
  ) %>%
  ggplot(aes(x = type, y = pan_predictions, fill = PFI)) +
  
  geom_boxplot(
    position = position_dodge(width = 0.8),
    outlier.alpha = 0.4
  ) +
  
  stat_compare_means(
    aes(group = PFI),
    method = "wilcox.test",     # or "t.test"
    label = "p.signif",         # stars: *, **, ***
    hide.ns = TRUE              # hide "ns"
  ) +
  
  labs(
    x = "Cancer type",
    y = "Pan prediction",
    fill = "Outcome"
  ) +
  
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid.major.x = element_blank()
  )
ggsave("LP_Boxplots.pdf", width = 12, height = 6)
```


##multivariate cox
```{r}
get_vars_cox <- function(cox_model)
{
  beta <- coef(cox_model)
ci   <- confint(cox_model)
pval <- summary(cox_model)$coefficients[, "Pr(>|z|)"]

  variable = names(beta)
  estimate = round(exp(beta), 2)
  CI_low   = round(exp(ci[, 1]), 2)
  CI_high  = round(exp(ci[, 2]), 2)
 # p_value = formatC(pval, format = "e", digits = 2)
  #p_value = round(pval, 3)
  p_value <- ifelse(pval < 0.001, "p < 0.001",
                         paste("p = ", formatC(pval, format = "f", digits = 3), sep = ""))

  estimates <- c()
  for(var in variable)
  {
    if (length(variable) == 1)
    {
          value <- paste(estimate[var], " [", (CI_low), " - ", (CI_high), "]", " ", p_value, sep = "")
    }
    else
    {
          value <- paste(estimate[var], " [", (CI_low[var]), " - ", (CI_high[var]), "]", " ", p_value[var], sep = "")
    }
    #print(var)

    #print(value)
    estimates <- c(estimates, value)
  }
  return (estimates)
}
estimates <- get_vars_cox(individual_cox_2)
print(estimates)
```


```{r}
pre_mice_rt <- pre_mice %>% filter(radiotherapy == "yes")
complete_cases_rt <- pre_mice_rt #%>% drop_na(any_of(c("type", "merged_stage", "age", "gender")))
complete_cases_rt <- complete_cases_rt %>% filter(PFI.time != 0)

for (typeC in examine_types)
{
  typeData <- complete_cases_rt %>% filter(type == typeC)
  four_fifths <- sum(typeData$PFI) * 4/5
  epp10 = 10
  epp20 = 20
  degrees_freedom_10 = four_fifths/epp10
  degrees_freedom_20 = four_fifths/epp20
  print(paste(typeC, four_fifths, degrees_freedom_10, degrees_freedom_20))
}
pca_numbers_10 <- c(7, 5, 10, 12, 7, 10, 4, 6)
pca_numbers_20 <- c(3, 2, 5, 6, 4, 5, 2, 3)
```


```{r}
generate_PC_formula <- function(numPCA)
{
  numbers <- 1:(numPCA-1)
  PCs <- rep("PC", numPCA-1)
  finalPC <- paste("PC", numPCA, sep = "")
  return(as.formula(paste("Surv(PFI.time, PFI) ~", paste(PCs, numbers, " + ", sep = "", collapse = ""), finalPC)))
}
validate_and_return_c <- function(coxmodel, nboot)
{
  tryCatch(val <- validate(coxmodel, B = nboot, conf.int = 0.95), error = function(e){return(NA)})
  tryCatch(val['Dxy', 'index.corrected']/2, error = function(e){return(NA)})
  
  estimate <- val['Dxy', 'index.corrected']/2 + 1/2
  low <- val['Dxy', 'Lower']/2 + 1/2
  high <- val['Dxy', 'Upper']/2 + 1/2
  
  estimate <- round(estimate, 2)
  low <- round(low, 2)
  high <- round(high, 2)
  
  toreturn <- paste(estimate, " [", low, " - ", high, "]", sep = "")
  
  return(toreturn)
}

validate_and_return_r <- function(coxmodel, nboot)
{
  tryCatch(val <- validate(coxmodel, B = nboot, conf.int = 0.95), error = function(e){return(NA)})
  tryCatch(val['R2', 'index.corrected']/2, error = function(e){return(NA)})
  
  estimate <- val['R2', 'index.corrected']
  low <- val['R2', 'Lower']
  high <- val['R2', 'Upper']
  
  estimate <- round(estimate, 2)
  low <- round(low, 2)
  high <- round(high, 2)
  
  toreturn <- paste(estimate, " [", low, " - ", high, "]", sep = "")
  
  return(toreturn)
}
```


```{r warning = F}

library(ggsurvfit)
library(survminer)
library(caret)

nfolds = 5
nboot = 500
nrepeats = 10

examine_types <-  c("BRCA", "CESC", "HNSC", "LGG", "LUAD","SKCM", "THCA", "UCEC")

pre_mice_rt <- pre_mice %>% filter(radiotherapy == "yes")
complete_cases_rt <- pre_mice_rt 
complete_cases_rt <- complete_cases_rt %>% filter(PFI.time != 0)

complete_cases_rt <- complete_cases_rt %>% mutate(type = as.character(type))
complete_cases_rt <- complete_cases_rt %>% mutate(type = ifelse(type %in% examine_types, type, "OTHER"))
complete_cases_rt <- complete_cases_rt %>% mutate(type = as.factor(type))

plots = list()

estimates_df_1 <- data.frame(matrix(ncol = 1, nrow = 8)) #ncol = 1, nrow = 8
rownames(estimates_df_1) = examine_types
colnames(estimates_df_1) = "pan-cancer"

estimates_df_6 <- data.frame(matrix(ncol = 1, nrow = 8)) #ncol = 1, nrow = 8
rownames(estimates_df_6) = examine_types
colnames(estimates_df_6) = "per-cancer-100"

estimates_df_7 <- data.frame(matrix(ncol = 1, nrow = 8)) #ncol = 1, nrow = 8
rownames(estimates_df_7) = examine_types
colnames(estimates_df_7) = "per-cancer-10EPP"

estimates_df_8 <- data.frame(matrix(ncol = 1, nrow = 8)) #ncol = 1, nrow = 8
rownames(estimates_df_8) = examine_types
colnames(estimates_df_8) = "per-cancer-20EPP"

estimates_df_2 <- data.frame(matrix(ncol = 6, nrow = 8)) #ncol = 6, nrow = 8)
rownames(estimates_df_2) = examine_types
colnames(estimates_df_2) = c("pan-cancer", "age", "stage2", "stage3", "stage4", "male")

estimates_df_3 <- data.frame(matrix(ncol = 7, nrow = 8)) #ncol = 6, nrow = 8)
rownames(estimates_df_3) = examine_types
colnames(estimates_df_3) = c("pan-cancer", "per-cancer", "age", "stage2", "stage3", "stage4", "male")

estimates_df_4 <- data.frame(matrix(ncol = 7, nrow = 8)) #ncol = 6, nrow = 8)
rownames(estimates_df_4) = examine_types
colnames(estimates_df_4) = c("pan-cancer", "per-cancer", "age", "stage2", "stage3", "stage4", "male")

estimates_df_5 <- data.frame(matrix(ncol = 7, nrow = 8)) #ncol = 6, nrow = 8)
rownames(estimates_df_5) = examine_types
colnames(estimates_df_5) = c("pan-cancer", "per-cancer", "age", "stage2", "stage3", "stage4", "male")

total_df <- data.frame(matrix(ncol = 6, nrow = 8))
colnames(total_df) <- c("clinical", "pan", "per", "clinical_pan", "clinical_per", "clinical_pan_per")
rownames(total_df) <- examine_types

r2_df <- data.frame(matrix(ncol = 6, nrow = 8))
colnames(r2_df) <- c("clinical", "pan", "per", "clinical_pan", "clinical_per", "clinical_pan_per")
rownames(r2_df) <- examine_types

typeCounter = 1
for (typeC in examine_types)
{
    print(typeC)
  type_data <- complete_cases_rt %>% filter(type == typeC) #cases of the evaluation type
  non_type_data <- complete_cases_rt %>% filter(type != typeC) #all other cases
  
  #pan-cancer predictions
  formula_all <- as.formula(paste("Surv(PFI.time, PFI) ~", paste(all_genes,  collapse = " + "), "+ type"))
  cox_pan_cancer <- coxph(formula_all, data = non_type_data)
    
  test_pan_predictions <- predict(cox_pan_cancer, type = "lp", newdata = type_data)
    test_pan_predictions <- as.numeric(test_pan_predictions)
    type_data$pan_predictions <- test_pan_predictions
   
    sd_pan <- sd(type_data$pan_predictions)
    type_data <- type_data %>% mutate(pan_predictions = pan_predictions/sd_pan) #STANDARDIZE FOR INTERPRETATION
    
  individual_cox <- coxph(Surv(PFI.time, PFI) ~ pan_predictions, data = type_data)

  #per-cancer, full model
  repeat_predictions <- data.frame()
  for (repI in 1:nrepeats)
  {
      folds <- createFolds(type_data$PFI, list = T, k = nfolds) #create folds of per-cancer data
      
      hold_predictions <- data.frame()
  
      pca_explained_10 <- c()
      pca_explained_20 <- c()
  
     for (i in 1:nfolds) #do per-cancer cross validated predictions
  {
    train_fold <- type_data[-as.numeric(unlist(folds[i])),] 
     test_fold <- type_data[as.numeric(unlist(folds[i])),]
    
     #get basic model
      formula_all <- as.formula(paste("Surv(PFI.time, PFI) ~", paste(all_genes,  collapse = " + ")))
     cox_per_cancer <- coxph(formula_all, data = train_fold)
     
     per_predictions <- predict(cox_per_cancer, type = "lp", newdata = test_fold)
     test_fold$per_predictions = per_predictions
     
     #PCA models
    lengthPCA_10 = pca_numbers_10[typeCounter]
    lengthPCA_20 = pca_numbers_20[typeCounter]
    pca_formula_10 <- generate_PC_formula(lengthPCA_10)
    pca_formula_20 <- generate_PC_formula(lengthPCA_20)
     
    train_fold_selected <- train_fold %>% dplyr::select(all_of(all_genes))
    test_fold_selected <- test_fold %>% dplyr::select(all_of(all_genes))
      
    #transformation calculated from train only
    pca_preprocess_params <- preProcess(train_fold_selected, method = c("center", "scale"))
    #apply transformation to both train and test
    train_scaled <- predict(pca_preprocess_params, train_fold_selected)
    test_scaled <- predict(pca_preprocess_params, test_fold_selected)

    #fit PCA model
    pca_model <- prcomp(train_scaled, center = FALSE, scale. = FALSE) #no need to transform already done
    s <- summary(pca_model)
    pca_explained_10 <- c(pca_explained_10, s$importance['Cumulative Proportion', lengthPCA_10]) #cumulative explained
    pca_explained_20 <- c(pca_explained_20, s$importance['Cumulative Proportion', lengthPCA_20]) #cumulative explained
    
    #get PCA space for train and test
    train_pca_coords <- predict(pca_model, newdata = train_scaled)
    test_pca_coords <- predict(pca_model, newdata = test_scaled)
  
    #grab lengthPCA components
    final_train_data <- train_pca_coords[, 1:lengthPCA_10] #this is the bigger number params (vs 20)
    final_test_data <- test_pca_coords[, 1:lengthPCA_10]
    train_fold <- cbind(train_fold, final_train_data)
    test_fold <- cbind(test_fold, final_test_data)

    cox_pca_10 <- coxph(pca_formula_10, data = train_fold) #ALT 1 - NO ADJUSTMENT
    test_predictions_10 <- predict(cox_pca_10, newdata = test_fold)

    cox_pca_20 <- coxph(pca_formula_20, data = train_fold) #ALT 1 - NO ADJUSTMENT
    test_predictions_20 <- predict(cox_pca_20, newdata = test_fold)
    
    test_fold$per_predictions_2 <- test_predictions_10
    test_fold$per_predictions_3 <- test_predictions_20
  
    #merge in
    hold_predictions <- rbind(hold_predictions, test_fold)
    }
      
    hold_predictions <- hold_predictions %>% dplyr::select(case, per_predictions, per_predictions_2, per_predictions_3)
    repeat_predictions <- rbind(repeat_predictions, hold_predictions)
    # pca_explained_10 <- mean(pca_explained_10)
    # print(paste("PCA 10:", pca_explained_10))
    # pca_explained_20 <- mean(pca_explained_20)
    # print(paste("PCA 20:", pca_explained_20))
    
    
    
  }
   
    repeat_predictions <- repeat_predictions %>% group_by(case) %>% summarise(per_predictions = mean(per_predictions, na.rm = TRUE),
    per_predictions_2 = mean(per_predictions_2, na.rm = TRUE),
    per_predictions_3 = mean(per_predictions_3, na.rm = TRUE))
  
  type_data <- left_join(type_data, hold_predictions, by = join_by(case))
  
  #standardize per
      sd_per1 <- sd(type_data$per_predictions)
      sd_per2 <- sd(type_data$per_predictions_2)
      sd_per3 <- sd(type_data$per_predictions_3)
    type_data <- type_data %>% mutate(per_predictions = per_predictions/sd_per1, per_predictions_2 = per_predictions_2/sd_per2, per_predictions_3 = per_predictions_3/sd_per3) #STANDARDIZE FOR INTERPRETATION
  
  #for per-cancer evaluation, need to drop
    type_data <- type_data %>% drop_na(any_of(c("type", "merged_stage", "age", "gender"))) 

   
      if (typeC == "BRCA" || typeC == "CESC" || typeC == "UCEC") #sex-specific cancers, no gender var
     {
          individual_cox_2 <- coxph(Surv(PFI.time, PFI) ~ pan_predictions + age + merged_stage, data = type_data)
          individual_cox_3 <- coxph(Surv(PFI.time, PFI) ~ pan_predictions + per_predictions + age + merged_stage, data = type_data)
                    individual_cox_4 <- coxph(Surv(PFI.time, PFI) ~ pan_predictions + per_predictions_2 + age + merged_stage, data = type_data)
                     individual_cox_5 <- coxph(Surv(PFI.time, PFI) ~ pan_predictions + per_predictions_3 + age + merged_stage, data = type_data)
      }else
      {
        individual_cox_2 <- coxph(Surv(PFI.time, PFI) ~ pan_predictions + age + merged_stage + gender, data = type_data)
         individual_cox_3 <- coxph(Surv(PFI.time, PFI) ~ pan_predictions + per_predictions + age + merged_stage + gender, data = type_data)
                  individual_cox_4 <- coxph(Surv(PFI.time, PFI) ~ pan_predictions + per_predictions_2 + age + merged_stage + gender, data = type_data)
                  individual_cox_5 <- coxph(Surv(PFI.time, PFI) ~ pan_predictions + per_predictions_3 + age + merged_stage + gender, data = type_data)
      }
    
      individual_cox_6 <- coxph(Surv(PFI.time, PFI) ~ per_predictions, data = type_data)
      individual_cox_7 <- coxph(Surv(PFI.time, PFI) ~ per_predictions_2, data = type_data)
      individual_cox_8 <- coxph(Surv(PFI.time, PFI) ~ per_predictions_3, data = type_data)
  
  estimates_1 <- get_vars_cox(individual_cox)
  estimates_2 <- get_vars_cox(individual_cox_2)
  estimates_3 <- get_vars_cox(individual_cox_3)
   estimates_4 <- get_vars_cox(individual_cox_4)
   estimates_5 <- get_vars_cox(individual_cox_5)
   
     estimates_6 <- get_vars_cox(individual_cox_6)
       estimates_7 <- get_vars_cox(individual_cox_7)
         estimates_8 <- get_vars_cox(individual_cox_8)
  
  if (length(estimates_2) == 5) #if no gender var
  {
    estimates_2 <- c(estimates_2, NA)
  }
  
    if (length(estimates_3) == 6) #if no gender var
  {
    estimates_3 <- c(estimates_3, NA)
    }
   
       if (length(estimates_4) == 6) #if no gender var
  {
    estimates_4 <- c(estimates_4, NA)
  }
  
          if (length(estimates_5) == 6) #if no gender var
  {
    estimates_5 <- c(estimates_5, NA)
  }
  
  
  #print(estimates_3)
  estimates_df_1[typeCounter,] = estimates_1
   estimates_df_2[typeCounter,] = estimates_2
   estimates_df_3[typeCounter,] = estimates_3
  estimates_df_4[typeCounter,] = estimates_4
   estimates_df_5[typeCounter,] = estimates_5
    estimates_df_6[typeCounter,] = estimates_6
    estimates_df_7[typeCounter,] = estimates_7
    estimates_df_8[typeCounter,] = estimates_8
    

    type_data <- droplevels(type_data) #gets rid of missing category problem

    if (typeC == "BRCA" || typeC == "CESC" || typeC == "UCEC") #sex-specific cancers, no gender var
     {
    
          clinical <- cph(Surv(PFI.time, PFI) ~ merged_stage + age, data = type_data, x = T, y = T)
          pan <- cph(Surv(PFI.time, PFI) ~ pan_predictions, data = type_data, x = T, y = T)
          per <- cph(Surv(PFI.time, PFI) ~ per_predictions_2, data = type_data, x = T, y = T)
          clinical_pan <- cph(Surv(PFI.time, PFI) ~ merged_stage + age + pan_predictions, data = type_data, x = T, y = T)
          clinical_per <- cph(Surv(PFI.time, PFI) ~ merged_stage + age + per_predictions_2, data = type_data, x = T, y = T)
          clinical_pan_per <- cph(Surv(PFI.time, PFI) ~ merged_stage + age + per_predictions_2 + pan_predictions, data = type_data, x = T, y = T)
      }else
      {
        
          clinical <- cph(Surv(PFI.time, PFI) ~ merged_stage + age + gender, data = type_data, x = T, y = T)
          pan <- cph(Surv(PFI.time, PFI) ~ pan_predictions, data = type_data, x = T, y = T)
          per <- cph(Surv(PFI.time, PFI) ~ per_predictions_2, data = type_data, x = T, y = T)
          clinical_pan <- cph(Surv(PFI.time, PFI) ~ merged_stage + age + gender + pan_predictions, data = type_data, x = T, y = T)
          clinical_per <- cph(Surv(PFI.time, PFI) ~ merged_stage + age + gender + per_predictions_2, data = type_data, x = T, y = T)
          clinical_pan_per <- cph(Surv(PFI.time, PFI) ~ merged_stage + age + gender + per_predictions_2 + pan_predictions, data = type_data, x = T, y = T)
         
      }
    
      result_list_c <- c(validate_and_return_c(clinical, nboot), 
                       validate_and_return_c(pan, nboot), 
                       validate_and_return_c(per, nboot), 
                       validate_and_return_c(clinical_pan, nboot), 
                       validate_and_return_c(clinical_per, nboot), 
                       validate_and_return_c(clinical_pan_per, nboot))
      
      total_df[typeCounter, ] = result_list_c
      
            result_list_r <- c(validate_and_return_r(clinical, nboot), 
                       validate_and_return_r(pan, nboot), 
                       validate_and_return_r(per, nboot), 
                       validate_and_return_r(clinical_pan, nboot), 
                       validate_and_return_r(clinical_per, nboot), 
                       validate_and_return_r(clinical_pan_per, nboot))
      
      r2_df[typeCounter, ] = result_list_r
      
      
  typeCounter = typeCounter + 1
}
```



```{r}
write.csv(estimates_df_1, "estimates_1.csv")
write.csv(estimates_df_2, "estimates_2.csv")
write.csv(estimates_df_3, "estimates_3.csv")
write.csv(estimates_df_4, "estimates_4.csv")
write.csv(estimates_df_5, "estimates_5.csv")
write.csv(estimates_df_6, "estimates_6.csv")
write.csv(estimates_df_7, "estimates_7.csv")
write.csv(estimates_df_8, "estimates_8.csv")
write.csv(total_df, "total_df.csv")
write.csv(r2_df, "r2_df.csv")
```


```{r}
estimates_df_6 
estimates_df_7
estimates_df_8
estimates_df_3
estimates_df_4
estimates_df_5
```

```{r}
df6 <- as.data.frame(t(estimates_df_6))
df7 <- as.data.frame(t(estimates_df_7))
df8 <- as.data.frame(t(estimates_df_8))
df3 <- as.data.frame(t(estimates_df_3))
df4 <- as.data.frame(t(estimates_df_4))
df5 <- as.data.frame(t(estimates_df_5))
all_df <- rbind(df6, df3, df7, df4, df8, df5)
View(all_df)
write.csv(all_df, "per_cancer_results.csv")
```

```{r}

```



