---
title: "Pan_Cancer_Merged_Analysis"
output: pdf_document
date: "2025-06-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#REPLACE WITH YOUR DIRECTORY
knitr::opts_knit$set(root.dir = "C:\\Users\\harry\\OneDrive\\Documents\\phd\\tcga_pan_cancer")
```

#setup
```{r}
library(tidyverse)
library(mice)
library(survival)
library(rms)
library(survminer)
```

if doing gene set analysis, run this chunk first, then run the gene section, then unload these libraries (they interfere with tidyverse)
```{r}
library(singscore)
library(GSEABase)
library(msigdb)
library(ExperimentHub)
```

#process mRNA data
##load hallmark genes
```{r}
eh = ExperimentHub()
query(eh , 'msigdb')
msigdb.hs = getMsigdb(org = 'hs', id = 'SYM', version = '7.4')
msigdb.hs
listCollections(msigdb.hs)
hallmarks <- subsetCollection(msigdb.hs, 'h')
```

##apply singscore algorithm in batches
(this requires your mRNA files downloaded locally)
```{r warning = FALSE}
fraction_dataframe <- data.frame()
setwd("./m11_zip")
folders <- list.files()
length(folders)
index = 1
for (folder in folders)
{
  print(index)
  setwd(folder)
  patient_id <- folder
  tsv_file <- list.files(pattern = "\\.tsv$", full.names = TRUE, recursive = TRUE)[1]
  test_tsv <- read.table(tsv_file, sep = '\t', header = TRUE)
  
  length_tsv <- length(test_tsv$gene_name)
  test_tsv_mod <- test_tsv %>% dplyr::select(gene_name, tpm_unstranded)
  test_tsv_mod <- test_tsv_mod[5:length_tsv, ]
  test_tsv_mod <- test_tsv_mod %>% distinct(gene_name, .keep_all = TRUE)
  
  test_tsv_1col <- data.frame(test_tsv_mod$tpm_unstranded)
  rownames(test_tsv_1col) <- test_tsv_mod$gene_name
  test_tsv_1col_filter <- test_tsv_1col %>% filter(test_tsv_mod.tpm_unstranded != 0)
  rankData_test <- rankGenes(test_tsv_1col_filter)
  
  names <- c()
  scores_dir <-  c()
  scores_undir <-  c()
  dispersions_dir <- c()
  dispersions_undir <- c()
  
  scoredf_test <- simpleScore(rankData_test, h1data)
  scoredf_test
  
  for (hallmarkNumber in 1:50)
  {
    h1 <- hallmarks[hallmarkNumber] #check the actual sets we want 
    h1data <- h1[[1]] #this is a gene set
    name <- GSEABase::setName(h1data)
    score_dir <- simpleScore(rankData_test, h1data, knownDirection = TRUE)$TotalScore
    score_undir <- simpleScore(rankData_test, h1data, knownDirection = FALSE)$TotalScore
    dispersion_dir <- simpleScore(rankData_test, h1data, knownDirection = TRUE)$TotalDispersion
    dispersion_undir <- simpleScore(rankData_test, h1data, knownDirection = FALSE)$TotalDispersion
    names <- c(names, name)
    scores_dir <- c(scores_dir, score_dir)
    scores_undir <- c(scores_undir, score_undir)
    dispersions_dir <- c(dispersions_dir, dispersion_dir)
    dispersions_undir <- c(dispersions_undir, dispersion_undir)
  }
  

  #add to combined dataframe 
  data_list <- c(patient_id)
  names(data_list) <- "Patient_ID"
  for (i in 1:50) {
    data_list[paste0(names[i], "_Score_Dir")] <- scores_dir[i]
    data_list[paste0(names[i], "_Score_Undir")] <- scores_undir[i]
    data_list[paste0(names[i], "_Dispersion_Dir")] <- dispersions_dir[i]
    data_list[paste0(names[i], "_Dispersion_Undir")] <- dispersions_undir[i]
  }
  
  df <- as.data.frame(t(data_list), stringsAsFactors = FALSE)
  df[-1] <- lapply(df[-1], as.numeric)
  fraction_dataframe <- rbind(fraction_dataframe, df)
  
  setwd("..")
  
  index = index + 1
}

write.csv(fraction_dataframe, "m11_zip_extracted.csv")
```

##merge batches
```{r}
big_df <- data.frame()
setwd("./outputs")
files = list.files()
for (file in files)
{
  temp_csv <- read.csv(file)
  big_df <- rbind(big_df, temp_csv)
}

write.csv(big_df, "full_gene_data.csv")
```

##add patient IDs
```{r}
file_mappings <- read.csv("json_file_mappings.csv")
file_mappings <- file_mappings %>% rename(case_id = patients)
file_mappings <- file_mappings %>% mutate(case_id = substr(case_id, start = 1, stop = 12))

big_df <- read.csv("full_gene_data.csv")
big_df <- big_df[, -c(1,2)]
big_df <- big_df %>% rename(files = Patient_ID)

json_merged <- left_join(big_df, file_mappings, by = "file_ids")

write.csv(json_merged, "gene_data_with_ids.csv")
```


##replace above code for RSI calculation

Gene expressions for the 10 genes rank-ordered, with lowest expression ranked 1.

then apply: 
RSI = 
- 0.0098009 * AR 
+ 0.0128283 * JUN
+ 0.0254552 * STAT1
- 0.0017589 * PRKCB 
- 0.0038171 * RELA
+ 0.1070213 * ABL1
- 0.0002509 * SUMO1
- 0.0092431 * CDK1 
- 0.0204469 * HDAC1 
- 0.0441683 * IRF1

```{r warning = FALSE}
fraction_dataframe <- data.frame()
setwd("./m12_zip")
folders <- list.files()
length(folders)
index = 1
for (folder in folders)
{
  if(index %% 100 == 0){print(index)}
  setwd(folder)
  patient_id <- folder
  tsv_file <- list.files(pattern = "\\.tsv$", full.names = TRUE, recursive = TRUE)[1]
  test_tsv <- read.table(tsv_file, sep = '\t', header = TRUE)
  
  length_tsv <- length(test_tsv$gene_name)
  test_tsv_mod <- test_tsv %>% dplyr::select(gene_name, tpm_unstranded)
  test_tsv_mod <- test_tsv_mod[5:length_tsv, ]
  test_tsv_mod <- test_tsv_mod %>% distinct(gene_name, .keep_all = TRUE)
  
  test_tsv_1col <- data.frame(test_tsv_mod$tpm_unstranded)
  rownames(test_tsv_1col) <- test_tsv_mod$gene_name
  test_tsv_1col_filter <- test_tsv_1col %>% filter(test_tsv_mod.tpm_unstranded != 0)
  
  
  rsi_genes <- c("AR", "JUN", "STAT1", "PRKCB", "RELA", "ABL1", "SUMO1", "CDK1", "HDAC1", "IRF1")
  rsi_coefs <- c(-0.0098009, 0.0128283, 0.0254552, -0.0017589, -0.0038171, 0.1070213, -0.0002509, -0.0092431, -0.0204469, 0.0441683)
  seq_10 <- seq(1,10)
  exprs <- c()
  for(gene in rsi_genes)
  {
    expr <- test_tsv_1col[gene,]
    exprs <- c(exprs, expr)
  }
  
  comb <- data.frame(rsi_genes, rsi_coefs, exprs)
  comb <- comb %>% arrange(exprs)
  comb <- cbind(comb, seq_10)
  comb <- comb %>% mutate(val = seq_10 * rsi_coefs)
  rsi <- sum(comb$val)
  df <- data.frame("Patient_ID" = patient_id, "RSI" = rsi)
  
  fraction_dataframe <- rbind(fraction_dataframe, df)
  
  setwd("..")
  
  index = index + 1
}
```

#create data files for further processing
##clinical TCGA
###load and select
```{r}
mmc_data <- read.csv("mmc1.csv")
mmc_small <- mmc_data %>% select(bcr_patient_barcode,type, age_at_initial_pathologic_diagnosis, gender, ajcc_pathologic_tumor_stage, clinical_stage, histological_type, histological_grade, initial_pathologic_dx_year, PFI, PFI.time, OS, OS.time)
#for LAML, no PFI, so replace with OS
mmc_small <- mmc_small %>% mutate(PFI = ifelse(type == "LAML", OS, PFI))
mmc_small <- mmc_small %>% select(-c(OS, OS.time))
mmc_mut <- mmc_small
mmc_mut <- mmc_mut %>% rename(age = age_at_initial_pathologic_diagnosis, path_stage = ajcc_pathologic_tumor_stage, year = initial_pathologic_dx_year, case = bcr_patient_barcode)
```

###parse missingness
```{r}
mmc_mut <- mmc_mut %>% mutate(
    type = ifelse(type == "#N/A", NA, type), 
    age = ifelse(age == "#N/A", NA, age), 
    gender = ifelse(gender == "#N/A", NA, gender), 
    path_stage = ifelse(path_stage == "#N/A", NA, path_stage), 
    clinical_stage = ifelse(clinical_stage == "#N/A", NA, clinical_stage), 
    histological_type = ifelse(histological_type == "#N/A", NA, histological_type), 
    histological_grade = ifelse(histological_grade == "#N/A", NA, histological_grade), 
    year = ifelse(year == "#N/A", NA, year), 
    PFI = ifelse(PFI == "#N/A", NA, PFI), 
    PFI.time = ifelse(PFI.time == "#N/A", NA, PFI.time)
)
mmc_mut <- mmc_mut %>% mutate(
  path_stage = ifelse(path_stage %in% c("[Not Available]", "[Not Applicable]", "[Discrepancy]", "[Unknown]", "Stage X"), NA, path_stage), 
  clinical_stage = ifelse(clinical_stage %in% c("[Not Available]", "[Not Applicable]", "[Discrepancy]", "[Unknown]"), NA, clinical_stage)
)
```

###create 4 stage categories
```{r}

stage1path <- c("I/II NOS", "IS", "Stage I", "Stage IA", "Stage IB")
stage2path <- c("Stage II", "Stage IIA", "Stage IIB", "Stage IIC")
stage3path <- c("Stage III", "Stage IIIA", "Stage IIIB", "Stage IIIC")
stage4path <- c("Stage IV", "Stage IVA", "Stage IVB", "Stage IVC")

stage1clin <- c("I", "Stage I", "Stage IA", "Stage IA1", "Stage IA2", "Stage IB", "Stage IB1", "Stage IB2", "Stage IC", "Stage IS")
stage2clin <- c("IIa", "IIb", "Stage II", "Stage IIA", "Stage IIA1", "Stage IIA2", "Stage IIB", "Stage IIC")
stage3clin <- c("III", "Stage III", "Stage IIIA", "Stage IIIB", "Stage IIIC", "Stage IIIC1", "Stage IIIC2")
stage4clin <- c("IVa", "IVb", "Stage IV", "Stage IVA", "Stage IVB", "Stage IVC")


mmc_mut <- mmc_mut %>% mutate( 
  path_stage = ifelse(path_stage %in% stage1path, "Stage 1", 
                      ifelse(path_stage %in% stage2path, "Stage 2", 
                             ifelse(path_stage %in% stage3path, "Stage 3", 
                                    ifelse(path_stage %in% stage4path, "Stage 4", NA)))))
mmc_mut <- mmc_mut %>% mutate( 
  clinical_stage = ifelse(clinical_stage %in% stage1clin, "Stage 1", 
                      ifelse(clinical_stage %in% stage2clin, "Stage 2", 
                             ifelse(clinical_stage %in% stage3clin, "Stage 3", 
                                    ifelse(clinical_stage %in% stage4clin, "Stage 4", NA)))))
mmc_mut <- mmc_mut %>% mutate(merged_stage = ifelse(is.na(path_stage), clinical_stage, path_stage))                        
```

##save
```{r}
mmc_premice <- mmc_mut
mmc_premice_save <- write.csv(mmc_premice, "mmc_premice_save.csv")
```

##radiotherapy treatment info
###RT dose information
```{r}
setwd("./all_bcr_biotabs")
getwd()
folders <- list.files()
folders <- folders[1:(length(folders)-1)] #dont go into the manifest
full_radio_df <- data.frame()

all_names <- c()

i = 0
for(folder in folders)
{
  folderpath <- file.path(".", folder)
 # print(folderpath)
  setwd(folderpath)
  files <- list.files()
  for(file in files)
  {
    if(grepl("radiation", file, fixed = TRUE)){
      print(file)
      tempdf <- read_tsv(file, show_col_types = FALSE)
      tempdf <- tempdf[-c(1,2),] #drop the first 2 rows (they're not actually data)
      #print(head(tempdf))
      #print(colnames(tempdf))
      
      if("measure_of_response" %in% colnames(tempdf))
      {
        tempdf <- tempdf %>% rename(treatment_best_response = measure_of_response)
      }
      if("regimen_indication" %in% colnames(tempdf))
      {
        tempdf <- tempdf %>% rename(therapy_regimen = regimen_indication)
      }
      if("regimen_indication_notes" %in% colnames(tempdf))
      {
        tempdf <- tempdf %>% rename(therapy_regimen_other = regimen_indication_notes)
      }
      if("days_to_radiation_therapy_start" %in% colnames(tempdf))
      {
        tempdf <- tempdf %>% rename(radiation_therapy_started_days_to = days_to_radiation_therapy_start)
      }
      if("days_to_radiation_therapy_end" %in% colnames(tempdf))
      {
        tempdf <- tempdf %>% rename(radiation_therapy_ended_days_to = days_to_radiation_therapy_end)
      }
      if("radiation_type_notes" %in% colnames(tempdf))
      {
        tempdf <- tempdf %>% rename(radiation_type_other = radiation_type_notes)
      }

      all_names <- c(all_names, colnames(tempdf))
      
      full_radio_df <- rbind(full_radio_df, tempdf)
    }
  }
  #print(files)
  setwd("..")
}
table(all_names)
```

###clean courses of treatment
```{r}
#select vars
full_radio_df_small <- full_radio_df %>% select(bcr_patient_barcode, radiation_total_dose, radiation_therapy_type, radiation_adjuvant_units, radiation_adjuvant_fractions_total)
full_radio_df_small <- full_radio_df_small %>% rename(case = bcr_patient_barcode)
#make numeric
full_radio_df_small <- full_radio_df_small %>% mutate(radiation_adjuvant_fractions_total = as.numeric(radiation_adjuvant_fractions_total))
full_radio_df_small <- full_radio_df_small %>% mutate(radiation_total_dose = as.numeric(radiation_total_dose))
#clean units
na_units <- c("[Discrepancy]", "[Not Available]")
full_radio_df_small <- full_radio_df_small %>% mutate(radiation_adjuvant_units = ifelse(radiation_adjuvant_units %in% na_units, NA, radiation_adjuvant_units))
#clean course
na_course <- c("[Discrepancy]", "[Not Available]", "[Unknown]", "OTHER")
full_radio_df_small <- full_radio_df_small %>% mutate(radiation_therapy_type = ifelse(radiation_therapy_type %in% na_course, NA, radiation_therapy_type))
full_radio_df_small <- full_radio_df_small %>% mutate(radiation_therapy_type = ifelse(radiation_therapy_type == "EXTERNAL BEAM", "External", radiation_therapy_type))
full_radio_df_small <- full_radio_df_small %>% mutate(radiation_therapy_type = ifelse(radiation_therapy_type == "IMPLANTS", "Internal", radiation_therapy_type))
full_radio_df_small <- full_radio_df_small %>% mutate(radiation_therapy_type = ifelse(radiation_therapy_type == "RADIOISOTOPE", "Systemic", radiation_therapy_type))
full_radio_df_small <- full_radio_df_small %>% mutate(radiation_therapy_type = ifelse(radiation_therapy_type == "COMBINATION", "Combined", radiation_therapy_type))
```

Based on (royal college radiologists), the following values were deemed impossible:
  Individual course of treatment with over 50 fractions
  Individual course of treatment with over 100 Gy, 300 mCi
  
```{r}
maxDose = 100
maxFrac = 50
mCiDose = 300

full_radio_df_small <- full_radio_df_small %>% mutate(converted_dose = 
                                                  ifelse(radiation_adjuvant_units == "cGy", radiation_total_dose/100, radiation_total_dose))
full_radio_df_small <- full_radio_df_small %>% mutate(converted_dose = ifelse(radiation_adjuvant_units == "mCi", radiation_total_dose,  ifelse(converted_dose > maxDose , converted_dose/100, converted_dose))) %>% #if not mCi, say >100 Gy is impossible 
mutate(converted_dose = ifelse(converted_dose > mCiDose, NA, converted_dose)) #if mCi, say > 300 mCi is impossible

full_radio_df_small <- full_radio_df_small %>% mutate(radiation_adjuvant_fractions_total = ifelse(radiation_adjuvant_fractions_total > maxFrac, NA, radiation_adjuvant_fractions_total))
```

```{r}
courses_merged <- full_radio_df_small %>% group_by(case) %>% 
  summarise(radiation_adjuvant_fractions_total = sum(radiation_adjuvant_fractions_total), 
            radiation_therapy_type = paste0(radiation_therapy_type, collapse = ""), 
            converted_dose = sum(converted_dose)) %>% 
  ungroup()
```

###fix radiotherapy type
```{r}
table(courses_merged$radiation_therapy_type)
#Combined External Internal Systemic
i = 1
for (item in courses_merged$radiation_therapy_type)
{
  combined <- grepl("Combined", item)
  external <- grepl("External", item)
  internal <- grepl("Internal", item)
  systemic <- grepl("Systemic", item)
  amount <- sum(c(combined, external, internal, systemic))
  if (amount == 0)
  {
    courses_merged[i, "radiation_therapy_type"] = NA
  }
  else if (amount > 1)
  {
    courses_merged[i, "radiation_therapy_type"] = "Combined"
  }
  else if (combined == T)
  {
    courses_merged[i, "radiation_therapy_type"] = "Combined"
  }
  else if (external == T)
  {
    courses_merged[i, "radiation_therapy_type"] = "External"
  }
    else if (internal == T)
  {
    courses_merged[i, "radiation_therapy_type"] = "Internal"
    }
   else if (systemic == T)
  {
    courses_merged[i, "radiation_therapy_type"] = "Systemic"
   }
  else
  {
    print("ERROR")
  }
  i = i + 1
}
table(courses_merged$radiation_therapy_type, useNA = "always")
```


```{r}
write.csv(courses_merged, "rtx_info.csv")
```

#create dataframe
##genes
GSEA scoring extraction is in gsea_testing.rmd 
```{r}
gene_data_ids <- read.csv("gene_data_with_ids.csv")
manifest_full <- read_tsv("tumor_type_manifest.txt")
gene_ids_tumour <- gene_data_ids %>% filter(file_ids %in% manifest_full$id)
gene_ids_tumour <- gene_ids_tumour %>% select(-c(X, file_ids, ids))
```

##clinical treatment
```{r}
full_clinical_data <- read.csv("full_clinical_file.csv")
full_clinical_data <- full_clinical_data %>% rename(case = case_submitter_id)

full_clinical_radio <- full_clinical_data %>% filter(treatment_type == "Radiation Therapy, NOS")
full_clinical_radio <- full_clinical_radio %>% select(case, treatment_or_therapy)
full_clinical_radio <- full_clinical_radio %>% rename(case_id = case, radiotherapy = treatment_or_therapy)

full_clinical_chemo <- full_clinical_data %>% filter(treatment_type == "Pharmaceutical Therapy, NOS")
full_clinical_chemo <- full_clinical_chemo %>% select(case, treatment_or_therapy)
full_clinical_chemo <- full_clinical_chemo %>% rename(case_id = case, chemotherapy = treatment_or_therapy)

full_clinical_merged <- left_join(full_clinical_radio, full_clinical_chemo, by = join_by(case_id))

gene_data_merged <- left_join(full_clinical_merged, gene_ids_tumour, by = join_by(case_id))
```

##process
merge gene data across multiple samples
```{r}
gene_data_all <- gene_data_merged %>% rename(case = case_id)
gene_data_all <- gene_data_all %>% group_by(case) %>% 
  summarise(
    radiotherapy = first(radiotherapy),
    chemotherapy = first(chemotherapy),
    across(where(is.numeric), mean, na.rm = TRUE)) #am averaging all TUMOUR samples here

#divide by standard deviation and center
gene_data_all <- gene_data_all %>% mutate(across(where(is.numeric), ~scale(., center = T, scale = T)))

gene_data_all <- gene_data_all %>% mutate(radiotherapy = ifelse(radiotherapy == "not reported", NA, radiotherapy), chemotherapy = ifelse(chemotherapy == "not reported", NA, chemotherapy))
```

##clinical outcomes
```{r}
mmc_premice <- read.csv("mmc_premice_save.csv")
mmc_premice <- mmc_premice %>% select(-c(X))

pre_mice_dataframe <- left_join(gene_data_all, mmc_premice, by = join_by(case))

#exclude missing outcomes
pre_mice_dataframe <- pre_mice_dataframe %>% filter(!is.na(PFI)) %>% filter(!is.na(PFI.time))

#exclude missing genes
pre_mice_dataframe <- pre_mice_dataframe %>% filter(!is.na(HALLMARK_TNFA_SIGNALING_VIA_NFKB_Score_Dir))

```

##process

```{r}
#process grade
hist_na <- c("[Discrepancy]", "[Not Available]", "[Unknown]", "GX", "GB")
table(pre_mice_dataframe$histological_grade)
pre_mice_dataframe <- pre_mice_dataframe %>% mutate(histological_grade = ifelse(histological_grade %in% hist_na, NA, ifelse(histological_grade == "Low Grade", "G1", ifelse(histological_grade == "High Grade", "G3", histological_grade))))
#Low and High refer only to BLCA - Low grade is the same as grade 1. High grade is the same as grade 3.

#change types
pre_mice_dataframe <- pre_mice_dataframe %>% mutate(type = as.factor(type), age = as.numeric(age), gender = as.factor(gender), path_stage = as.factor(path_stage), clinical_stage = as.factor(clinical_stage), histological_grade = as.factor(histological_grade), year = as.numeric(year), PFI = as.numeric(PFI), PFI.time = as.numeric(PFI.time), merged_stage = as.factor(merged_stage))

#to include LGG, use grade instead of stage
pre_mice_complete <- pre_mice_dataframe %>% mutate(histological_grade = ifelse(histological_grade == "G1", "Stage 1", ifelse(histological_grade == "G2", "Stage 2", ifelse(histological_grade == "G3", "Stage 3", ifelse(histological_grade == "G4", "Stage 4", NA)))))
pre_mice_complete <- pre_mice_complete %>% mutate(merged_stage = as.character(merged_stage))
pre_mice_complete <- pre_mice_complete %>% mutate(merged_stage = ifelse(type == "LGG", histological_grade, merged_stage))
pre_mice_complete <- pre_mice_complete %>% mutate(merged_stage = as.factor(merged_stage))

#clean 
pre_mice_complete <- pre_mice_complete %>% select(-c(clinical_stage, path_stage))
pre_mice_complete <- pre_mice_complete %>% select(!contains("Undir"))
```

##radiotherapy treatment info
```{r}
rtx_info <- read.csv("rtx_info.csv")
rtx_info <- rtx_info %>% select(-c(X))
pre_mice_complete <- left_join(pre_mice_complete, rtx_info, by = join_by(case))
```

##purity
```{r}
tumor_purity <- read.csv("tumor_purity.csv")
lengthT <- length(tumor_purity$X)
tumor_purity2 <- tumor_purity[4:lengthT,1:7]
tumor_purity3 <- tumor_purity2 %>% mutate(case = tumor_purity2[,1], purity = ifelse(X.5 == "NaN", NA, X.5) )
tumor_purity3 <- tumor_purity3 %>% mutate(case = substr(case, start = 1, stop = 12))
tumor_purity3 <- tumor_purity3 %>% select(case, purity)
tumor_purity3 <- tumor_purity3 %>% mutate(purity = as.numeric(purity))
tumor_purity5 <- tumor_purity3[!duplicated(tumor_purity3$case),]
pre_mice_complete <- left_join(pre_mice_complete, tumor_purity5, by = join_by(case))
```

##mutational burden
```{r}
mutational_burden <- read.csv("mutational_burden.csv")
mutational_burden <- mutational_burden %>% dplyr::select("Patient.ID", "TMB") 
mutational_burden <- mutational_burden %>% dplyr::rename(case = Patient.ID)
pre_mice_complete <- left_join(pre_mice_complete, mutational_burden, by = join_by(case))
```

##write and read (fixes data types)
```{r}
write.csv(pre_mice_complete, "pre_mice_complete_may_22.csv")
pre_mice <- read.csv("pre_mice_complete_may_22.csv")
pre_mice <- pre_mice %>% select(-c(X, X.1))
pre_mice <- pre_mice %>% mutate(type = as.factor(type), age = as.numeric(age), gender = as.factor(gender), histological_grade = as.factor(histological_grade), year = as.numeric(year), PFI = as.numeric(PFI), PFI.time = as.numeric(PFI.time), merged_stage = as.factor(merged_stage))
```

##process
```{r}
pre_mice <- pre_mice %>% mutate(radiotherapy = as.factor(radiotherapy), chemotherapy = as.factor(chemotherapy), radiation_therapy_type = as.factor(radiation_therapy_type))

#z-score normalize converted dose, fractions and center age, year
pre_mice_norm <- pre_mice %>% group_by(type) %>% mutate(converted_dose = ((converted_dose - mean(converted_dose, na.rm = T))/sd(converted_dose, na.rm = T))) %>% ungroup()
pre_mice_norm <- pre_mice_norm %>% group_by(type) %>% mutate(radiation_adjuvant_fractions_total = ((radiation_adjuvant_fractions_total - mean(radiation_adjuvant_fractions_total, na.rm = T))/sd(radiation_adjuvant_fractions_total, na.rm = T))) %>% ungroup()
pre_mice_norm <- pre_mice_norm %>% mutate(age = age - mean(age, na.rm = T)) %>% mutate(year = year - mean(year, na.rm = T))
hist(pre_mice_norm$converted_dose)
hist(pre_mice_norm$radiation_adjuvant_fractions_total)
pre_mice_norm <- pre_mice_norm %>% mutate(TMB = log(TMB))
```

#add RSI
```{r}
rsi_data <- read.csv("gene_data_with_ids_rsi.csv")
rsi_data <- rsi_data %>% dplyr::filter(files %in% manifest_full$id)
rsi_data <- rsi_data %>% dplyr::select(-c(X, X.1, files, ids))

rsi_data <- rsi_data %>% dplyr::rename(case = case_id)
rsi_data <- rsi_data %>% group_by(case) %>% 
  summarise(
    across(where(is.numeric), mean, na.rm = TRUE)) #am averaging all TUMOUR samples here

pre_mice_norm <- left_join(pre_mice_norm, rsi_data, by = join_by(case))
```

#mice
```{r}
mice_df <- mice(pre_mice_norm, m = 50)
```

##make complete dataframe for evaluation
```{r}
complete_cases_rt <- pre_mice_norm %>% filter(radiotherapy == "yes" & !is.na(merged_stage) & !is.na(age) & !is.na(gender))
```

#explore and tables
##table 1
```{r}
my.render.cont <- function(x) {
    with(stats.apply.rounding(stats.default(x), digits=2), c("",
        "Mean (SD)"=sprintf("%s (%s)", MEAN, SD)))
}

my.render.cat <- function(x) {
    c("", sapply(stats.default(x), function(y) with(y,
        sprintf("%d (%0.f%%)", FREQ, PCT))))
}
pvalue <- function(x, ...) {
    # Construct vectors of data y, and groups (strata) g
    y <- unlist(x)
    g <- factor(rep(1:length(x), times=sapply(x, length)))
    if (is.numeric(y)) {
        # For numeric variables, perform a standard 2-sample t-test
        p <- t.test(y ~ g)$p.value
    } else {
        # For categorical variables, perform a chi-squared test of independence
        p <- chisq.test(table(y, g))$p.value
    }
    # Format the p-value, using an HTML entity for the less-than sign.
    # The initial empty string places the output on the line below the variable label.
    c("", sub("<", "&lt;", format.pval(p, digits=3, eps=0.001)))
}
```

https://cran.r-project.org/web/packages/table1/vignettes/table1-latex.pdf

pre_mice_dataframe
rt_df
pre_mice_complete

```{r}
library(table1)

table1_df <- pre_mice %>% filter(is.na(radiotherapy) | radiotherapy == "yes")

table1_df <- table1_df %>% rename(fractions = radiation_adjuvant_fractions_total, rt_type = radiation_therapy_type, dose = converted_dose)

t1_recurrence <- table1(~ radiotherapy + merged_stage + histological_grade +  age  + gender + year + chemotherapy + dose + rt_type + fractions + purity + TMB | PFI, data=table1_df, render.continuous = my.render.cont, render.categorical = my.render.cat)
#print(t1_recurrence)

#library(flextable)
#x1 <- t1flex(t1_recurrence)
#print(xtable(x1))


# x1 <- t1kable(t1_recurrence)
# print(x1)
x1 <- t1kable(t1_recurrence, format = "latex", longtable = T)
print(x1)
writeLines(x1, 'complete_patients_dose_new.tex')
```

##flowchart
11228     
  -148 missing all clinical data other than cancer type
  = 11080
outcomes
  -173 missing progression or progression time
  = 10907
genes
  -636 missing tumour gene expression data
  = 10271
rt
  -6004 not confirmed RT
  = 4267


```{r}
# Load the DiagrammeR package
library(DiagrammeR)

# Create the flowchart
g_flow <- DiagrammeR::grViz("
digraph flowchart {
  graph [layout = dot, rankdir = TB]

  # Main flow nodes
  node [shape = rectangle, fontname = Arial, fontsize = 12, style = filled, fillcolor = lightblue]
  A [label = '11228 in TCGA clinical files']
  B [label = '11080 remain']
  C [label = '10907 remain']
  D [label = '10271 remain']
  E [label = '4267 remain']

  # Exclusion nodes
  node [fillcolor = lightpink]
  Ex1 [label = '148 excluded:\\nMissing all clinical data']
  Ex2 [label = '173 excluded:\\nMissing event status or time']
  Ex3 [label = '636 excluded:\\nMissing gene expression data']
  Ex4 [label = '6004 excluded:\\nConfirmed no radiotherapy']
  
  # Edges
  edge [color = black, arrowhead = vee]

  # Main vertical flow
  A -> B
  B -> C
  C -> D
  D -> E

  # Horizontal branches for exclusions directly from vertical arrows
  edge [style = dashed]
  A -> Ex1 [constraint = false, tailport = s, headport = w]
  B -> Ex2 [constraint = false, tailport = s, headport = w]
  C -> Ex3 [constraint = false, tailport = s, headport = w]
  D -> Ex4 [constraint = false, tailport = s, headport = w]
  
    # Node alignment
  { rank = same; B; Ex1 }
  { rank = same; C; Ex2 }
  { rank = same; D; Ex3 }
  { rank = same; E; Ex4 }
  
}
")
g_flow
```

```{r}
library(DiagrammeR)
library(DiagrammeRsvg)
library(magrittr)
library(rsvg)

g_flow %>%
    export_svg %>% charToRaw %>% rsvg_pdf(".//supplementary_S1_flowchart.pdf")
```


##breakdown by type
```{r}

library(forcats)
figure1_df <- complete_cases_rt
# Plot with reordered cancer types
ggplot(figure1_df, aes(x = fct_reorder(type, -table(type)[type]), fill = factor(PFI, levels = c(1, 0)))) +
  geom_bar() +
  labs(x = "Cancer Type", y = "Number of Patients", fill = "PFI") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 12, angle = 60, hjust = 1), plot.title = element_text(size = 16, face = "bold", hjust = 0.5), axis.title = element_text(size = 14, face = "bold"), legend.title = element_text(size = 14, face = "bold"),   legend.text = element_text(size = 12)) + 
  ggtitle("Patients by Cancer Type and Progression")  + scale_fill_manual(values = c("lightblue", "pink"))

ggsave("Selected_by_types.pdf", width = 10, height = 7)
```

#analysis
##redundancy
```{r}
formula <- as.formula(paste("~", paste(c(colnames(pre_mice_norm)[4:103]),  collapse = " + ")))
r3 <- redun(formula, nk = 0, data = pre_mice_norm, r2 = 0.6, type = "adjusted")
```

```{r}
library(kableExtra)
x <- r3$rsquared[1:62]
names <- names(x)
i = 1
for (name in names)
{
  nch <- nchar(names[i])-4
  names[i] = substr(names[i], 10, nch)
  i = i+1
}
rsq <- round(as.numeric(x), 2)
df_print <- data.frame(names = names, rsq = rsq)
x1 <- kable(df_print, format = "latex", booktabs = T, longtable = T)
writeLines(x1, 'withheld_genes_new.tex')
```

```{r}
x <- r3$Out[1:62]
all_genes <- colnames(pre_mice_norm)[4:103]
select_genes_2 <- all_genes[!all_genes %in% x]
select_genes_2
```

##gene set estimates
```{r}
formula_cox <- as.formula(paste("Surv(PFI.time, PFI) ~", paste(select_genes_2,  collapse = " + "), "+ TMB + purity + chemotherapy + merged_stage + age + gender + radiation_adjuvant_fractions_total + radiation_therapy_type + converted_dose + year + histological_grade + type"))
PFI_mi_calc = with(data=mice_df, coxph(Surv(PFI.time, PFI) ~ HALLMARK_HYPOXIA_Dispersion_Dir + HALLMARK_CHOLESTEROL_HOMEOSTASIS_Dispersion_Dir + 
    HALLMARK_MITOTIC_SPINDLE_Dispersion_Dir + HALLMARK_WNT_BETA_CATENIN_SIGNALING_Dispersion_Dir + 
    HALLMARK_TGF_BETA_SIGNALING_Dispersion_Dir + HALLMARK_DNA_REPAIR_Dispersion_Dir + 
    HALLMARK_NOTCH_SIGNALING_Dispersion_Dir + HALLMARK_ADIPOGENESIS_Dispersion_Dir + 
    HALLMARK_ESTROGEN_RESPONSE_EARLY_Dispersion_Dir + HALLMARK_ANDROGEN_RESPONSE_Dispersion_Dir + 
    HALLMARK_MYOGENESIS_Dispersion_Dir + HALLMARK_PROTEIN_SECRETION_Dispersion_Dir + 
    HALLMARK_INTERFERON_ALPHA_RESPONSE_Dispersion_Dir + HALLMARK_APICAL_JUNCTION_Dispersion_Dir + 
    HALLMARK_APICAL_SURFACE_Dispersion_Dir + HALLMARK_HEDGEHOG_SIGNALING_Dispersion_Dir + 
    HALLMARK_COMPLEMENT_Dispersion_Dir + HALLMARK_UNFOLDED_PROTEIN_RESPONSE_Dispersion_Dir + 
    HALLMARK_PI3K_AKT_MTOR_SIGNALING_Dispersion_Dir + HALLMARK_MYC_TARGETS_V2_Dispersion_Dir + 
    HALLMARK_OXIDATIVE_PHOSPHORYLATION_Dispersion_Dir + HALLMARK_REACTIVE_OXYGEN_SPECIES_PATHWAY_Score_Dir + 
    HALLMARK_REACTIVE_OXYGEN_SPECIES_PATHWAY_Dispersion_Dir + 
    HALLMARK_P53_PATHWAY_Dispersion_Dir + HALLMARK_UV_RESPONSE_UP_Dispersion_Dir + 
    HALLMARK_UV_RESPONSE_DN_Dispersion_Dir + HALLMARK_ANGIOGENESIS_Dispersion_Dir + 
    HALLMARK_HEME_METABOLISM_Dispersion_Dir + HALLMARK_BILE_ACID_METABOLISM_Dispersion_Dir + 
    HALLMARK_PEROXISOME_Score_Dir + HALLMARK_PEROXISOME_Dispersion_Dir + 
    HALLMARK_SPERMATOGENESIS_Score_Dir + HALLMARK_SPERMATOGENESIS_Dispersion_Dir + 
    HALLMARK_KRAS_SIGNALING_UP_Dispersion_Dir + HALLMARK_KRAS_SIGNALING_DN_Score_Dir + 
    HALLMARK_KRAS_SIGNALING_DN_Dispersion_Dir + HALLMARK_PANCREAS_BETA_CELLS_Score_Dir + 
    HALLMARK_PANCREAS_BETA_CELLS_Dispersion_Dir + TMB + purity + 
    chemotherapy + merged_stage + age + gender + radiation_adjuvant_fractions_total + 
    radiation_therapy_type + converted_dose + year + histological_grade + 
    type))
#summary(PFI_mi_calc)

pooled_results <- pool(PFI_mi_calc)

summary_results <- summary(pooled_results)
summary_results$HR <- exp(summary_results$estimate)  # Convert log HR to HR
summary_results$HR_lower <- exp(summary_results$estimate - 1.96 * summary_results$std.error)  # Lower bound
summary_results$HR_upper <- exp(summary_results$estimate + 1.96 * summary_results$std.error)  # Upper bound

library(knitr)
# Display results
summary_results[, c("term", "HR", "HR_lower", "HR_upper")]

# Select columns and format
table_data <- summary_results[, c("term", "HR", "HR_lower", "HR_upper", "p.value")]
names(table_data) <- c("Variable", "HR", "Lower CI", "Upper CI", "p-value")

# Create a formatted table with kable

kable(
  table_data,
  digits = c(NA, 2, 2, 2, 4),
  caption = "Pooled Cox Regression Results PFI"
)
```

```{r}
df_print <- table_data %>% filter(Variable %in% select_genes_2) 
names(df_print) <- c("Variable", "HR", "LowerCI", "UpperCI", "pvalue")
df_print <- df_print %>% arrange(pvalue)
df_print <- df_print %>% mutate(HR = round(HR, 2), LowerCI = round(LowerCI,2), UpperCI = round(UpperCI, 2), pvalue = round(pvalue, 2))
names(df_print) <- c("Variable", "HR", "Lower CI", "Upper CI", "p-value")
x1 <- kable(df_print, format = "latex", row.names = F, booktabs = T, longtable = T, caption = "Hazard ratios, 95% confidence intervals, and p-values for 38 included gene set parameters. Parameters are ordered by most significant p values.") 
writeLines(x1, 'estimates_genes.tex')
```

```{r}
ranges <- c()
IQRs <- c()
for (gene in select_genes_2)
{
  range <- range(pre_mice_norm[,gene])[2] - range(pre_mice_norm[,gene])[1]
  iqr <- IQR(unlist(pre_mice_norm[,gene]))
  
  print(gene)
  print(range)
  print(iqr)
 
  ranges <- c(ranges, range)
  IQRs <- c(IQRs, iqr)
}
mean(ranges)
mean(IQRs)

```


##main analysis

###main
```{r}
library(nricens)

ntype = 8 #FIX

total_df <- data.frame(matrix(ncol = 9, nrow = ntype))
colnames(total_df) <- c("HR", "HR_L", "HR_U", "C_F", "C_G",  "R2_F","R2_G", "Q_F", "Q_G")

#types with over 100 RT patients
types_over_one_hundred <- c("BRCA", "CESC", "HNSC", "LGG", "LUAD","SKCM", "THCA", "UCEC")

plots <- list()

rownames(total_df) <- types_over_one_hundred

t0_time <- 365*5
nimpute = mice_df$m

violin_df <- complete(mice_df, 1) %>% select(case, type, PFI)

typeI = 1
for (typeC in types_over_one_hundred)
{
  
  print(typeC)

  d0 <- complete(mice_df, 1) %>% filter(type == typeC)
  
  d0_cases_orig <- data.frame(d0$case) %>% rename(case = d0.case)
  d0_cases <- d0_cases_orig
  
  for (i in 1:nimpute)
  {
      d1 <- complete(mice_df, i) %>% filter(radiotherapy == "yes")
    
        d1 <- d1 %>%
        add_count(type) %>%  # adds a column `n` with counts per type
        mutate(type = if_else(n < 50, as.factor("other"), type)) %>%
        select(-n)  # optionally remove the count column
      
      d_withhold <- d1 %>% filter(type != typeC) 
      d_type <- d1 %>% filter(type == typeC)
     
         formula_outer <- as.formula(paste("Surv(PFI.time, PFI) ~", paste(select_genes_2,  collapse = " + "), "+ TMB + purity + chemotherapy + merged_stage + age + gender + radiation_adjuvant_fractions_total + radiation_therapy_type + converted_dose + year + histological_grade + type"))
        cox_clinicogenomic_withheld <- coxph(formula_outer, data = d_withhold)
       
        predictions_withheld_terms <- data.frame(predict(cox_clinicogenomic_withheld, type = "terms", newdata = d_type))
      predictions_withheld_terms <- predictions_withheld_terms[,1:length(select_genes_2)]
       rowsums_terms <- rowSums(predictions_withheld_terms)
      d_type$predictions = rowsums_terms
    
      d_type_predictions <- d_type %>% select(case, predictions)
      d0_cases <- left_join(d0_cases, d_type_predictions, by = join_by(case))
  }
  
  to_merge <- d0_cases[,2:nimpute+1]
  to_merge_predictions <- rowSums(to_merge)
  d0_cases_orig$predictions = to_merge_predictions

  d2 <- complete_cases_rt %>% filter(type == typeC)
  if(typeC == "LGG")
  {
    d2 <- droplevels(d2)
  }
  d2 <- left_join(d2, d0_cases_orig, by = join_by(case))
  d2_iqr <- IQR(d2$predictions)
  d2 <- d2 %>% mutate(predictions = predictions/d2_iqr)

  if (typeC == "BRCA" || typeC == "CESC" || typeC == "UCEC") #sex-specific cancers
        { 
      f <- coxph(Surv(PFI.time, PFI) ~ merged_stage + age, data = d2) 
      g <- coxph(Surv(PFI.time, PFI) ~ predictions + merged_stage + age, data = d2)   

      f2 <- cph(Surv(PFI.time, PFI) ~ merged_stage + age, data = d2, x = T, y = T, surv = T) 
      g2 <- cph(Surv(PFI.time, PFI) ~ predictions + merged_stage + age, data = d2, x = T, y = T, surv = T) 
  }
  else
  {
      f <- coxph(Surv(PFI.time, PFI) ~ merged_stage + age + gender, data = d2) 
      g <- coxph(Surv(PFI.time, PFI) ~ predictions + merged_stage + age + gender, data = d2) 
      
      f2 <- cph(Surv(PFI.time, PFI) ~ merged_stage + age + gender, data = d2, x = T, y = T, surv = T) 
      g2 <- cph(Surv(PFI.time, PFI) ~ predictions + merged_stage + age + gender, data = d2, x = T, y = T, surv = T) 
  }
  
  val_f <- validate(f2, method = "boot", B = 1000)
  c_f <- val_f['Dxy', 'index.corrected']/2 + 1/2
  r_f <- val_f['R2', 'index.corrected']
  q_f <- val_f['Q', 'index.corrected']
  
  val_g <- validate(g2, method = "boot", B = 1000)
  c_g <- val_g['Dxy', 'index.corrected']/2 + 1/2
  r_g <- val_g['R2', 'index.corrected']
  q_g <- val_g['Q', 'index.corrected']
  
    hr <- summary(g)$coefficients[1,2]
    hr_l <- exp(summary(g)$coefficients[1,1] - summary(g)$coefficients[1,3]*1.96) 
    hr_u <- exp(summary(g)$coefficients[1,1] + summary(g)$coefficients[1,3]*1.96) 
    
  
    #("HR", "HR_L", "HR_U", "C_F", "C_G",  "R2_F","R2_G", "Q_F", "Q_G")
  total_df[typeI, 1] = hr
  total_df[typeI, 2] = hr_l
  total_df[typeI, 3] = hr_u
  total_df[typeI, 4] = c_f
  total_df[typeI, 5] = c_g
  total_df[typeI, 6] = r_f
  total_df[typeI, 7] = r_g
  total_df[typeI, 8] = q_f
  total_df[typeI, 9] = q_g
  
     d3 <- d2 %>%
        mutate(genetic_score = cut(predictions,
                              breaks = quantile(predictions, probs = c(0, 1/2, 1), na.rm = TRUE),
                              labels = c("Lower Half", "Upper Half"),
                              include.lowest = TRUE))

      d3 <- d3 %>% mutate(genetic_score = as.factor(genetic_score))

      # Create the survival object
      S_PFI <- Surv(d3$PFI.time, d3$PFI)

      # Fit the Kaplan-Meier model using the cut scores
      PFI.kmfit <- survfit(S_PFI ~ genetic_score, data = d3, conf.int = 0.95)

      # Plot the survival curves
      g_plot <- ggsurvplot(fit = PFI.kmfit, # Kaplan-Meier survival model
                 palette = c("pink", "lightblue"), # Custom color palette for three groups
                 legend.title = "Genetic Risk Score", # Legend title
                 legend.labs = c("Lower Half", "Upper Half"), # Label lines
                 title = typeC,
                 xlab = "Time (days)", conf.int = TRUE, pval = TRUE, xlim = c(0,1825), break.time.by = 365
                 )
    
    plots[[typeI]] <- g_plot
  
     typeI = typeI + 1
     
     pred <- d2 %>% select(case, predictions)
     violin_df <- left_join(violin_df, pred, by = join_by(case))
}

#violin_df <- violin_df %>% filter(!is.na(predictions))
violin_df <- violin_df %>%
  mutate(predictions = coalesce(predictions.x, predictions.y, predictions.x.x, predictions.y.y, predictions.x.x.x, predictions.y.y.y, predictions.x.x.x.x, predictions.y.y.y.y)) %>%
  select(-c(predictions.x, predictions.y, predictions.x.x, predictions.y.y, predictions.x.x.x, predictions.y.y.y, predictions.x.x.x.x, predictions.y.y.y.y))
```

##plots

###violin

```{r}
violin_df2 <- violin_df %>% filter(!is.na(predictions))
violin_df2 <- violin_df2 %>% mutate(PFI = as.factor(PFI))

p <- ggplot(violin_df2, aes(x=type, y=predictions, fill=PFI)) +
  geom_violin(position = position_dodge(width = 0.8)) +
  geom_boxplot(width = 0.1, position = position_dodge(width = 0.8), outlier.shape = NA, color = "black", show.legend=F) +
  xlab("Cancer Type") +
  ylab("Linear Predictions") +
  labs(title = "Genomic Predictions by Cancer Type and Outcome") + 
  scale_fill_manual(values = c("pink", "lightblue"))

p

#ggsave("violins.pdf", width = 8, height = 4)
#ggsave("violins_adversarial.pdf", width = 8, height = 4)
#ggsave("violins_logistic.pdf", width = 8, height = 4)
ggsave("violins_everyone.pdf", width = 8, height = 4)
```

```{r}

ranges <- c()
iqrs <- c()

for (typeC in types_over_one_hundred)
{
  typeData <- violin_df2 %>% filter(type == typeC)
  print(typeC)
  print(fivenum(typeData$predictions))
  print(IQR(typeData$predictions))
  print(range(typeData$predictions)[2] - range(typeData$predictions)[1])
  
  ranges <- c(ranges, range(typeData$predictions)[2] - range(typeData$predictions)[1])
  iqrs <- c(iqrs, IQR(typeData$predictions))
}

print(mean(ranges))
print(mean(iqrs))
```


###KM
```{r}
myplot <- arrange_ggsurvplots(plots, ncol = 2, nrow = 4) 
#myplot %>% ggsave(device="pdf", filename="survplots.pdf", width = 12, height = 16, units = "in")
#myplot %>% ggsave(device="pdf", filename="survplots_adversarial.pdf", width = 12, height = 16, units = "in")
#myplot %>% ggsave(device="pdf", filename="survplots_logistic.pdf", width = 12, height = 16, units = "in")
myplot %>% ggsave(device="pdf", filename="survplots_everyone.pdf", width = 12, height = 16, units = "in")
```

###forest
```{r}
forest_data <- total_df[, c("HR", "HR_L", "HR_U")]
forest_data$term <- rownames(total_df)
forest_data %>%
  ggplot(aes(x=term, y=HR, ymin=HR_L, ymax=HR_U)) +
  geom_point(size = 2) + 
  geom_errorbar(color = "black", width = 0.2) +
  geom_hline(yintercept = 1, color = "black") + 
  scale_x_discrete(limits = rev(rownames(total_df))) + 
  coord_flip() +
  xlab(" ") +
   ylab("Hazard Ratio (95% Confidence Interval)") + 
  labs(title ="Per-Cancer Prognostic Score Hazard Ratios") +
   theme(
     plot.title = element_text(size = 14, face = "bold"),
     axis.text.x = element_text(size = 12),
     axis.text.y = element_text(size = 12)
     ) + theme_minimal() #+
# theme(axis.text.x = element_text(size = 12), axis.text.y = element_text(size = 12), plot.title = element_text(size = 16, face = "bold", hjust = 0.5), axis.title = element_text(size = 14, face = "bold")) 

#ggsave("forest.pdf", width = 8, height = 5)
#ggsave("forest_adversarial.pdf", width = 8, height = 5)
#ggsave("forest_logistic.pdf", width = 8, height = 5)
ggsave("forest_everyone.pdf", width = 8, height = 5)
```

```{r}
forest_data
```


###CRQ
```{r}
library(gridExtra)

c_df <- total_df %>% dplyr::select(c("C_F", "C_G"))
c_df <- c_df %>% mutate(types = rownames(c_df))
df_long <- c_df %>%
  pivot_longer(cols = c(C_F, C_G), names_to = "Group", values_to = "Value") 
# Create the grouped bar plot
c_plot <- ggplot(df_long, aes(x = types, y = Value, fill = Group)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +  # Grouped bars
  scale_fill_manual(values = c("C_F" = "pink", "C_G" = "lightblue"), labels = c("Clinical", "Clinical + Genomic")) +  # Custom colors
  labs(title = "C Statistics With and Without Prognostic Scores",
       x = "Cancer Type",
       y = "Corrected C Statistic",
       fill = "Model") +
 
  theme_minimal()+ 
  coord_cartesian(ylim = c(0.5,0.7)) #normal
  #coord_cartesian(ylim = c(0.4,0.6)) #adversarial

c_df <- total_df %>% dplyr::select(c("R2_F", "R2_G"))
c_df <- c_df %>% mutate(types = rownames(c_df))
df_long <- c_df %>%
  pivot_longer(cols = c(R2_F, R2_G), names_to = "Group", values_to = "Value") 
# Create the grouped bar plot
r_plot <- ggplot(df_long, aes(x = types, y = Value, fill = Group)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +  # Grouped bars
  scale_fill_manual(values = c("R2_F" = "pink", "R2_G" = "lightblue"), labels = c("Clinical", "Clinical + Genomic")) +  # Custom colors
  labs(title = expression(paste("R"^2," Values With and Without Prognostic Scores")),
       x = "Cancer Type",
       y = expression("Corrected R"^2),
       fill = "Model") +
  theme_minimal()+ 
  #coord_cartesian(ylim = c(-0.03,0.15)) #normal
  #coord_cartesian(ylim = c(-0.06,0.01)) #adversarial
# coord_cartesian(ylim = c(-0.04,0.15)) #logistic
   coord_cartesian(ylim = c(-0.04,0.10)) #all

c_df <- total_df %>% dplyr::select(c("Q_F", "Q_G"))
c_df <- c_df %>% mutate(types = rownames(c_df))
df_long <- c_df %>%
  pivot_longer(cols = c(Q_F, Q_G), names_to = "Group", values_to = "Value") 
# Create the grouped bar plot
q_plot <- ggplot(df_long, aes(x = types, y = Value, fill = Group)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +  # Grouped bars
  scale_fill_manual(values = c("Q_F" = "pink", "Q_G" = "lightblue"), labels = c("Clinical", "Clinical + Genomic")) +  # Custom colors
  labs(title = expression(paste("Q Values With and Without Prognostic Scores")),
       x = "Cancer Type",
       y = "Corrected Q",
       fill = "Model") +
  theme_minimal()+ 
 # coord_cartesian(ylim = c(-0.015,0.03)) #normal
  #coord_cartesian(ylim = c(-0.06,0.01)) #adversarial
 # coord_cartesian(ylim = c(-0.02,0.03)) #logistic
  coord_cartesian(ylim = c(-0.005,0.035)) #all
  
g <- grid.arrange(c_plot, r_plot, q_plot, nrow = 3)

#ggsave("C_R_Q_plot.pdf", g, width = 8, height = 11 )
#ggsave("C_R_Q_plot_adversarial.pdf", g, width = 8, height = 11 )
#ggsave("C_R_Q_plot_logistic.pdf", g, width = 8, height = 11 )
ggsave("C_R_Q_plot_everyone.pdf", g, width = 8, height = 11 )
```

```{r}
c_df <- total_df %>% dplyr::select(c("Q_F", "Q_G"))
difs <- c_df$Q_G - c_df$Q_F
sort(difs)
mean(difs)
sd(difs)

df_test <- data.frame(names = types_over_one_hundred, difs = difs)
print(df_test)
```


#sensitivity

```{r}
  outcomes <- pre_mice_norm %>% select(case, PFI.time, PFI)
  exposures <- pre_mice_norm %>% select(-c(PFI.time, PFI))
  scrambled_cases <- sample(pre_mice_norm$case)
  outcomes$case = scrambled_cases
  
  scrambled_pre_mice <- left_join(outcomes, exposures, by = join_by(case))
```

```{r}
mice_scrambled <- mice(scrambled_pre_mice, m = 5)
```

```{r}
complete_cases_scrambled <- scrambled_pre_mice %>% filter(radiotherapy == "yes" & !is.na(merged_stage) & !is.na(age) & !is.na(gender))
```


##adversarial control
```{r}
library(nricens)

ntype = 8 #FIX

total_df <- data.frame(matrix(ncol = 9, nrow = ntype))
colnames(total_df) <- c("HR", "HR_L", "HR_U", "C_F", "C_G",  "R2_F","R2_G", "Q_F", "Q_G")

#types with over 100 RT patients
types_over_one_hundred <- c("BRCA", "CESC", "HNSC", "LGG", "LUAD","SKCM", "THCA", "UCEC")

plots <- list()

rownames(total_df) <- types_over_one_hundred

t0_time <- 365*5
nimpute = mice_scrambled$m

violin_df <- complete(mice_scrambled, 1) %>% select(case, type, PFI)

typeI = 1
for (typeC in types_over_one_hundred)
{
  
  print(typeC)

  d0 <- complete(mice_scrambled, 1) %>% filter(type == typeC)
  
  d0_cases_orig <- data.frame(d0$case) %>% rename(case = d0.case)
  d0_cases <- d0_cases_orig
  
  for (i in 1:nimpute)
  {
      d1 <- complete(mice_scrambled, i) %>% filter(radiotherapy == "yes")
      
        d1 <- d1 %>%
        add_count(type) %>%  # adds a column `n` with counts per type
        mutate(type = if_else(n < 50, as.factor("other"), type)) %>%
        select(-n)  # optionally remove the count column
      
      d_withhold <- d1 %>% filter(type != typeC) 
      d_type <- d1 %>% filter(type == typeC)
     
         formula_outer <- as.formula(paste("Surv(PFI.time, PFI) ~", paste(select_genes_2,  collapse = " + "), "+ TMB + purity + chemotherapy + merged_stage + age + gender + radiation_adjuvant_fractions_total + radiation_therapy_type + converted_dose + year + histological_grade + type"))
        cox_clinicogenomic_withheld <- coxph(formula_outer, data = d_withhold)
       
        predictions_withheld_terms <- data.frame(predict(cox_clinicogenomic_withheld, type = "terms", newdata = d_type))
      predictions_withheld_terms <- predictions_withheld_terms[,1:length(select_genes_2)]
       rowsums_terms <- rowSums(predictions_withheld_terms)
      d_type$predictions = rowsums_terms
    
      d_type_predictions <- d_type %>% select(case, predictions)
      d0_cases <- left_join(d0_cases, d_type_predictions, by = join_by(case))
  }
  
  to_merge <- d0_cases[,2:nimpute+1]
  to_merge_predictions <- rowSums(to_merge)
  d0_cases_orig$predictions = to_merge_predictions

  d2 <- complete_cases_scrambled %>% filter(type == typeC)
  if(typeC == "LGG")
  {
    d2 <- droplevels(d2)
  }
  d2 <- left_join(d2, d0_cases_orig, by = join_by(case))
  d2_iqr <- IQR(d2$predictions)
  d2 <- d2 %>% mutate(predictions = predictions/d2_iqr)
  
  if (typeC == "BRCA" || typeC == "CESC" || typeC == "UCEC") #sex-specific cancers
        { 
      f <- coxph(Surv(PFI.time, PFI) ~ merged_stage + age, data = d2) 
      g <- coxph(Surv(PFI.time, PFI) ~ predictions + merged_stage + age, data = d2)   

      f2 <- cph(Surv(PFI.time, PFI) ~ merged_stage + age, data = d2, x = T, y = T, surv = T) 
      g2 <- cph(Surv(PFI.time, PFI) ~ predictions + merged_stage + age, data = d2, x = T, y = T, surv = T) 
  }
  else
  {
      f <- coxph(Surv(PFI.time, PFI) ~ merged_stage + age + gender, data = d2) 
      g <- coxph(Surv(PFI.time, PFI) ~ predictions + merged_stage + age + gender, data = d2) 
      
      f2 <- cph(Surv(PFI.time, PFI) ~ merged_stage + age + gender, data = d2, x = T, y = T, surv = T) 
      g2 <- cph(Surv(PFI.time, PFI) ~ predictions + merged_stage + age + gender, data = d2, x = T, y = T, surv = T) 
  }
  
  val_f <- validate(f2, method = "boot", B = 1000)
  c_f <- val_f['Dxy', 'index.corrected']/2 + 1/2
  r_f <- val_f['R2', 'index.corrected']
  q_f <- val_f['Q', 'index.corrected']
  
  val_g <- validate(g2, method = "boot", B = 1000)
  c_g <- val_g['Dxy', 'index.corrected']/2 + 1/2
  r_g <- val_g['R2', 'index.corrected']
  q_g <- val_g['Q', 'index.corrected']
  
    hr <- summary(g)$coefficients[1,2]
    hr_l <- exp(summary(g)$coefficients[1,1] - summary(g)$coefficients[1,3]*1.96) 
    hr_u <- exp(summary(g)$coefficients[1,1] + summary(g)$coefficients[1,3]*1.96) 
    
  
    #("HR", "HR_L", "HR_U", "C_F", "C_G",  "R2_F","R2_G", "Q_F", "Q_G")
  total_df[typeI, 1] = hr
  total_df[typeI, 2] = hr_l
  total_df[typeI, 3] = hr_u
  total_df[typeI, 4] = c_f
  total_df[typeI, 5] = c_g
  total_df[typeI, 6] = r_f
  total_df[typeI, 7] = r_g
  total_df[typeI, 8] = q_f
  total_df[typeI, 9] = q_g
  
     d3 <- d2 %>%
        mutate(genetic_score = cut(predictions,
                              breaks = quantile(predictions, probs = c(0, 1/2, 1), na.rm = TRUE),
                              labels = c("Lower Half", "Upper Half"),
                              include.lowest = TRUE))

      d3 <- d3 %>% mutate(genetic_score = as.factor(genetic_score))

      # Create the survival object
      S_PFI <- Surv(d3$PFI.time, d3$PFI)

      # Fit the Kaplan-Meier model using the cut scores
      PFI.kmfit <- survfit(S_PFI ~ genetic_score, data = d3, conf.int = 0.95)

      # Plot the survival curves
      g_plot <- ggsurvplot(fit = PFI.kmfit, # Kaplan-Meier survival model
                 palette = c("pink", "lightblue"), # Custom color palette for three groups
                 legend.title = "Genetic Risk Score", # Legend title
                 legend.labs = c("Lower Half", "Upper Half"), # Label lines
                 title = typeC,
                 xlab = "Time (days)", conf.int = TRUE, pval = TRUE, xlim = c(0,1825), break.time.by = 365
                 )
    
    plots[[typeI]] <- g_plot
  
     typeI = typeI + 1
     
     pred <- d2 %>% select(case, predictions)
     violin_df <- left_join(violin_df, pred, by = join_by(case))
}

#violin_df <- violin_df %>% filter(!is.na(predictions))
violin_df <- violin_df %>%
  mutate(predictions = coalesce(predictions.x, predictions.y, predictions.x.x, predictions.y.y, predictions.x.x.x, predictions.y.y.y, predictions.x.x.x.x, predictions.y.y.y.y)) %>%
  select(-c(predictions.x, predictions.y, predictions.x.x, predictions.y.y, predictions.x.x.x, predictions.y.y.y, predictions.x.x.x.x, predictions.y.y.y.y))
```

##non coxph outer layer example

```{r}
library(nricens)

ntype = 8 #FIX

total_df <- data.frame(matrix(ncol = 9, nrow = ntype))
colnames(total_df) <- c("HR", "HR_L", "HR_U", "C_F", "C_G",  "R2_F","R2_G", "Q_F", "Q_G")

#types with over 100 RT patients
types_over_one_hundred <- c("BRCA", "CESC", "HNSC", "LGG", "LUAD","SKCM", "THCA", "UCEC")

plots <- list()

rownames(total_df) <- types_over_one_hundred

t0_time <- 365*5
nimpute = mice_df$m

violin_df <- complete(mice_df, 1) %>% select(case, type, PFI)

typeI = 1
for (typeC in types_over_one_hundred)
{
  
  print(typeC)

  d0 <- complete(mice_df, 1) %>% filter(type == typeC)
  
  d0_cases_orig <- data.frame(d0$case) %>% rename(case = d0.case)
  d0_cases <- d0_cases_orig
  
  for (i in 1:nimpute)
  {
      d1 <- complete(mice_df, i) %>% filter(radiotherapy == "yes")
    
        d1 <- d1 %>%
        add_count(type) %>%  # adds a column `n` with counts per type
        mutate(type = if_else(n < 50, as.factor("other"), type)) %>%
        select(-n)  # optionally remove the count column
      
      d_withhold <- d1 %>% filter(type != typeC) 
      d_type <- d1 %>% filter(type == typeC)
      
         formula_outer <- as.formula(paste("PFI ~", paste(select_genes_2,  collapse = " + "), "+ TMB + purity + chemotherapy + merged_stage + age + gender + radiation_adjuvant_fractions_total + radiation_therapy_type + converted_dose + year + histological_grade"))
        glm_clinicogenomic_withheld <- glm(formula_outer, data = d_withhold, family = "binomial")
        predictions_withheld_terms <- data.frame(predict(glm_clinicogenomic_withheld, type = "terms", newdata = d_type))
      predictions_withheld_terms <- predictions_withheld_terms[,1:length(select_genes_2)]
       rowsums_terms <- rowSums(predictions_withheld_terms)
      d_type$predictions = rowsums_terms
    
      d_type_predictions <- d_type %>% select(case, predictions)
      d0_cases <- left_join(d0_cases, d_type_predictions, by = join_by(case))
  }
  
  to_merge <- d0_cases[,2:nimpute+1]
  to_merge_predictions <- rowSums(to_merge)
  d0_cases_orig$predictions = to_merge_predictions

  d2 <- complete_cases_rt %>% filter(type == typeC)
  if(typeC == "LGG")
  {
    d2 <- droplevels(d2)
  }
  d2 <- left_join(d2, d0_cases_orig, by = join_by(case))
  d2_iqr <- IQR(d2$predictions)
  d2 <- d2 %>% mutate(predictions = predictions/d2_iqr)
  
  if (typeC == "BRCA" || typeC == "CESC" || typeC == "UCEC") #sex-specific cancers
        { 
      f <- coxph(Surv(PFI.time, PFI) ~ merged_stage + age, data = d2) 
      g <- coxph(Surv(PFI.time, PFI) ~ predictions + merged_stage + age, data = d2)   

      f2 <- cph(Surv(PFI.time, PFI) ~ merged_stage + age, data = d2, x = T, y = T, surv = T) 
      g2 <- cph(Surv(PFI.time, PFI) ~ predictions + merged_stage + age, data = d2, x = T, y = T, surv = T) 
  }
  else
  {
      f <- coxph(Surv(PFI.time, PFI) ~ merged_stage + age + gender, data = d2) 
      g <- coxph(Surv(PFI.time, PFI) ~ predictions + merged_stage + age + gender, data = d2) 
      
      f2 <- cph(Surv(PFI.time, PFI) ~ merged_stage + age + gender, data = d2, x = T, y = T, surv = T) 
      g2 <- cph(Surv(PFI.time, PFI) ~ predictions + merged_stage + age + gender, data = d2, x = T, y = T, surv = T) 
  }
  
  val_f <- validate(f2, method = "boot", B = 1000)
  c_f <- val_f['Dxy', 'index.corrected']/2 + 1/2
  r_f <- val_f['R2', 'index.corrected']
  q_f <- val_f['Q', 'index.corrected']
  
  val_g <- validate(g2, method = "boot", B = 1000)
  c_g <- val_g['Dxy', 'index.corrected']/2 + 1/2
  r_g <- val_g['R2', 'index.corrected']
  q_g <- val_g['Q', 'index.corrected']
  
    hr <- summary(g)$coefficients[1,2]
    hr_l <- exp(summary(g)$coefficients[1,1] - summary(g)$coefficients[1,3]*1.96) 
    hr_u <- exp(summary(g)$coefficients[1,1] + summary(g)$coefficients[1,3]*1.96) 
    
  
    #("HR", "HR_L", "HR_U", "C_F", "C_G",  "R2_F","R2_G", "Q_F", "Q_G")
  total_df[typeI, 1] = hr
  total_df[typeI, 2] = hr_l
  total_df[typeI, 3] = hr_u
  total_df[typeI, 4] = c_f
  total_df[typeI, 5] = c_g
  total_df[typeI, 6] = r_f
  total_df[typeI, 7] = r_g
  total_df[typeI, 8] = q_f
  total_df[typeI, 9] = q_g
  
     d3 <- d2 %>%
        mutate(genetic_score = cut(predictions,
                              breaks = quantile(predictions, probs = c(0, 1/2, 1), na.rm = TRUE),
                              labels = c("Lower Half", "Upper Half"),
                              include.lowest = TRUE))

      d3 <- d3 %>% mutate(genetic_score = as.factor(genetic_score))

      # Create the survival object
      S_PFI <- Surv(d3$PFI.time, d3$PFI)

      # Fit the Kaplan-Meier model using the cut scores
      PFI.kmfit <- survfit(S_PFI ~ genetic_score, data = d3, conf.int = 0.95)

      # Plot the survival curves
      g_plot <- ggsurvplot(fit = PFI.kmfit, # Kaplan-Meier survival model
                 palette = c("pink", "lightblue"), # Custom color palette for three groups
                 legend.title = "Genetic Risk Score", # Legend title
                 legend.labs = c("Lower Half", "Upper Half"), # Label lines
                 title = typeC,
                 xlab = "Time (days)", conf.int = TRUE, pval = TRUE, xlim = c(0,1825), break.time.by = 365
                 )
    
    plots[[typeI]] <- g_plot
  
     typeI = typeI + 1
     
     pred <- d2 %>% select(case, predictions)
     violin_df <- left_join(violin_df, pred, by = join_by(case))
}

#violin_df <- violin_df %>% filter(!is.na(predictions))
violin_df <- violin_df %>%
  mutate(predictions = coalesce(predictions.x, predictions.y, predictions.x.x, predictions.y.y, predictions.x.x.x, predictions.y.y.y, predictions.x.x.x.x, predictions.y.y.y.y)) %>%
  select(-c(predictions.x, predictions.y, predictions.x.x, predictions.y.y, predictions.x.x.x, predictions.y.y.y, predictions.x.x.x.x, predictions.y.y.y.y))
```

##don't restrict to radiotherapy

```{r}
complete_cases_everyone <- pre_mice_norm %>% filter(!is.na(merged_stage) & !is.na(age) & !is.na(gender))
```


```{r}
library(nricens)

ntype = 8 #FIX

total_df <- data.frame(matrix(ncol = 9, nrow = ntype))
colnames(total_df) <- c("HR", "HR_L", "HR_U", "C_F", "C_G",  "R2_F","R2_G", "Q_F", "Q_G")

#types with over 100 RT patients
types_over_one_hundred <- c("BRCA", "CESC", "HNSC", "LGG", "LUAD","SKCM", "THCA", "UCEC")

plots <- list()

rownames(total_df) <- types_over_one_hundred

t0_time <- 365*5
nimpute = mice_df$m

violin_df <- complete(mice_df, 1) %>% select(case, type, PFI)

typeI = 1
for (typeC in types_over_one_hundred)
{
  
  print(typeC)

  d0 <- complete(mice_df, 1) %>% filter(type == typeC)
  
  d0_cases_orig <- data.frame(d0$case) %>% rename(case = d0.case)
  d0_cases <- d0_cases_orig
  
  for (i in 1:nimpute)
  {
      d1 <- complete(mice_df, i)
    
        d1 <- d1 %>%
        add_count(type) %>%  # adds a column `n` with counts per type
        mutate(type = if_else(n < 50, as.factor("other"), type)) %>%
        select(-n)  # optionally remove the count column
      
      d_withhold <- d1 %>% filter(type != typeC) 
      d_type <- d1 %>% filter(type == typeC)
     
         formula_outer <- as.formula(paste("Surv(PFI.time, PFI) ~", paste(select_genes_2,  collapse = " + "), "+ TMB + purity + chemotherapy + merged_stage + age + gender + radiation_adjuvant_fractions_total + radiation_therapy_type + converted_dose + year + histological_grade + type"))
        cox_clinicogenomic_withheld <- coxph(formula_outer, data = d_withhold)
       
        predictions_withheld_terms <- data.frame(predict(cox_clinicogenomic_withheld, type = "terms", newdata = d_type))
      predictions_withheld_terms <- predictions_withheld_terms[,1:length(select_genes_2)]
       rowsums_terms <- rowSums(predictions_withheld_terms)
      d_type$predictions = rowsums_terms
    
      d_type_predictions <- d_type %>% select(case, predictions)
      d0_cases <- left_join(d0_cases, d_type_predictions, by = join_by(case))
  }
  
  to_merge <- d0_cases[,2:nimpute+1]
  to_merge_predictions <- rowSums(to_merge)
  d0_cases_orig$predictions = to_merge_predictions

  d2 <- complete_cases_everyone %>% filter(type == typeC)
  if(typeC == "LGG")
  {
    d2 <- droplevels(d2)
  }
  d2 <- left_join(d2, d0_cases_orig, by = join_by(case))
  d2_iqr <- IQR(d2$predictions)
  d2 <- d2 %>% mutate(predictions = predictions/d2_iqr)

  if (typeC == "BRCA" || typeC == "CESC" || typeC == "UCEC") #sex-specific cancers
        { 
      f <- coxph(Surv(PFI.time, PFI) ~ merged_stage + age, data = d2) 
      g <- coxph(Surv(PFI.time, PFI) ~ predictions + merged_stage + age, data = d2)   

      f2 <- cph(Surv(PFI.time, PFI) ~ merged_stage + age, data = d2, x = T, y = T, surv = T) 
      g2 <- cph(Surv(PFI.time, PFI) ~ predictions + merged_stage + age, data = d2, x = T, y = T, surv = T) 
  }
  else
  {
      f <- coxph(Surv(PFI.time, PFI) ~ merged_stage + age + gender, data = d2) 
      g <- coxph(Surv(PFI.time, PFI) ~ predictions + merged_stage + age + gender, data = d2) 
      
      f2 <- cph(Surv(PFI.time, PFI) ~ merged_stage + age + gender, data = d2, x = T, y = T, surv = T) 
      g2 <- cph(Surv(PFI.time, PFI) ~ predictions + merged_stage + age + gender, data = d2, x = T, y = T, surv = T) 
  }
  
  val_f <- validate(f2, method = "boot", B = 1000)
  c_f <- val_f['Dxy', 'index.corrected']/2 + 1/2
  r_f <- val_f['R2', 'index.corrected']
  q_f <- val_f['Q', 'index.corrected']
  
  val_g <- validate(g2, method = "boot", B = 1000)
  c_g <- val_g['Dxy', 'index.corrected']/2 + 1/2
  r_g <- val_g['R2', 'index.corrected']
  q_g <- val_g['Q', 'index.corrected']
  
    hr <- summary(g)$coefficients[1,2]
    hr_l <- exp(summary(g)$coefficients[1,1] - summary(g)$coefficients[1,3]*1.96) 
    hr_u <- exp(summary(g)$coefficients[1,1] + summary(g)$coefficients[1,3]*1.96) 
    
  
    #("HR", "HR_L", "HR_U", "C_F", "C_G",  "R2_F","R2_G", "Q_F", "Q_G")
  total_df[typeI, 1] = hr
  total_df[typeI, 2] = hr_l
  total_df[typeI, 3] = hr_u
  total_df[typeI, 4] = c_f
  total_df[typeI, 5] = c_g
  total_df[typeI, 6] = r_f
  total_df[typeI, 7] = r_g
  total_df[typeI, 8] = q_f
  total_df[typeI, 9] = q_g
  
     d3 <- d2 %>%
        mutate(genetic_score = cut(predictions,
                              breaks = quantile(predictions, probs = c(0, 1/2, 1), na.rm = TRUE),
                              labels = c("Lower Half", "Upper Half"),
                              include.lowest = TRUE))

      d3 <- d3 %>% mutate(genetic_score = as.factor(genetic_score))

      # Create the survival object
      S_PFI <- Surv(d3$PFI.time, d3$PFI)

      # Fit the Kaplan-Meier model using the cut scores
      PFI.kmfit <- survfit(S_PFI ~ genetic_score, data = d3, conf.int = 0.95)

      # Plot the survival curves
      g_plot <- ggsurvplot(fit = PFI.kmfit, # Kaplan-Meier survival model
                 palette = c("pink", "lightblue"), # Custom color palette for three groups
                 legend.title = "Genetic Risk Score", # Legend title
                 legend.labs = c("Lower Half", "Upper Half"), # Label lines
                 title = typeC,
                 xlab = "Time (days)", conf.int = TRUE, pval = TRUE, xlim = c(0,1825), break.time.by = 365
                 )
    
    plots[[typeI]] <- g_plot
  
     typeI = typeI + 1
     
     pred <- d2 %>% select(case, predictions)
     violin_df <- left_join(violin_df, pred, by = join_by(case))
}

#violin_df <- violin_df %>% filter(!is.na(predictions))
violin_df <- violin_df %>%
  mutate(predictions = coalesce(predictions.x, predictions.y, predictions.x.x, predictions.y.y, predictions.x.x.x, predictions.y.y.y, predictions.x.x.x.x, predictions.y.y.y.y)) %>%
  select(-c(predictions.x, predictions.y, predictions.x.x, predictions.y.y, predictions.x.x.x, predictions.y.y.y, predictions.x.x.x.x, predictions.y.y.y.y))
```


##compare with RSI

```{r}
library(nricens)

ntype = 8 #FIX

total_df <- data.frame(matrix(ncol = 15, nrow = ntype))
colnames(total_df) <- c("HR", "HR_L", "HR_U", "HR_RSI", "HR_L_RSI", "HR_U_RSI", "C_F", "C_G", "C_H","R2_F","R2_G", "R2_H", "Q_F", "Q_G", "Q_H")

#types with over 100 RT patients
types_over_one_hundred <- c("BRCA", "CESC", "HNSC", "LGG", "LUAD","SKCM", "THCA", "UCEC")
#types_over_one_hundred = c("CESC")

plots <- list()

rownames(total_df) <- types_over_one_hundred

t0_time <- 365*5

nimpute = 5

violin_df <- complete(mice_df, 1) %>% dplyr::select(case, type, PFI)
violin_df_RSI <- complete(mice_df, 1) %>% dplyr::select(case, type, PFI)

typeI = 1
for (typeC in types_over_one_hundred)
{
  
   print(typeC)

  d0 <- complete(mice_df, 1) %>% filter(type == typeC)
  
  d0_cases_orig <- data.frame(d0$case) %>% dplyr::rename(case = d0.case)
  d0_cases <- d0_cases_orig
  
  for (i in 1:nimpute)
  {
      d1 <- complete(mice_df, i) %>% filter(radiotherapy == "yes")
    
        d1 <- d1 %>%
        add_count(type) %>%  # adds a column `n` with counts per type
        mutate(type = if_else(n < 50, as.factor("other"), type)) %>%
        dplyr::select(-n)  # optionally remove the count column
      
      d_withhold <- d1 %>% filter(type != typeC) 
      #print(sum(d_withhold$PFI))
      d_type <- d1 %>% filter(type == typeC)
     
         formula_outer <- as.formula(paste("Surv(PFI.time, PFI) ~", paste(select_genes_2,  collapse = " + "), "+ TMB + purity + chemotherapy + merged_stage + age + gender + radiation_adjuvant_fractions_total + radiation_therapy_type + converted_dose + year + histological_grade + type"))
        cox_clinicogenomic_withheld <- coxph(formula_outer, data = d_withhold)
       
        predictions_withheld_terms <- data.frame(predict(cox_clinicogenomic_withheld, type = "terms", newdata = d_type))
      predictions_withheld_terms <- predictions_withheld_terms[,1:length(select_genes_2)]
       rowsums_terms <- rowSums(predictions_withheld_terms)
      d_type$predictions = rowsums_terms
    
      d_type_predictions <- d_type %>% dplyr::select(case, predictions)
      d0_cases <- left_join(d0_cases, d_type_predictions, by = join_by(case))
  }
  
  to_merge <- d0_cases[,2:nimpute+1]
  to_merge_predictions <- rowSums(to_merge)
  d0_cases_orig$predictions = to_merge_predictions
  
  d2 <- complete_cases_rt %>% filter(type == typeC)
  if(typeC == "LGG")
  {
    d2 <- droplevels(d2)
  }
  d2 <- left_join(d2, d0_cases_orig, by = join_by(case))
  
  d2 <- d2 %>% mutate(RSI = RSI*-1)
  d2_iqr <- IQR(d2$predictions)
  d2 <- d2 %>% mutate(predictions = predictions/d2_iqr)
  d2_iqr_rsi <- IQR(d2$RSI)
  d2 <- d2 %>% mutate(RSI = RSI/d2_iqr_rsi)
  
  if (typeC == "BRCA" || typeC == "CESC" || typeC == "UCEC") #sex-specific cancers
        { 
      f <- coxph(Surv(PFI.time, PFI) ~ merged_stage + age, data = d2) 
      g <- coxph(Surv(PFI.time, PFI) ~ predictions + merged_stage + age, data = d2)   
      h <- coxph(Surv(PFI.time, PFI) ~ RSI + merged_stage + age, data = d2)   

      f2 <- cph(Surv(PFI.time, PFI) ~ merged_stage + age , data = d2, x = T, y = T, surv = T) 
      g2 <- cph(Surv(PFI.time, PFI) ~ predictions + merged_stage + age, data = d2, x = T, y = T, surv = T) 
      h2 <- cph(Surv(PFI.time, PFI) ~ RSI + merged_stage + age, data = d2, x = T, y = T, surv = T) 
  }
  else
  {
      f <- coxph(Surv(PFI.time, PFI) ~ merged_stage + age + gender , data = d2) 
      g <- coxph(Surv(PFI.time, PFI) ~ predictions + merged_stage + age + gender, data = d2) 
      h <- coxph(Surv(PFI.time, PFI) ~ RSI + merged_stage + age + gender, data = d2) 
      
      f2 <- cph(Surv(PFI.time, PFI) ~ merged_stage + age + gender , data = d2, x = T, y = T, surv = T) 
      g2 <- cph(Surv(PFI.time, PFI) ~ predictions + merged_stage + age + gender, data = d2, x = T, y = T, surv = T) 
      h2 <- cph(Surv(PFI.time, PFI) ~ RSI + merged_stage + age + gender, data = d2, x = T, y = T, surv = T) 
  }
  
  val_f <- validate(f2, method = "boot", B = 1000)
  c_f <- val_f['Dxy', 'index.corrected']/2 + 1/2
  r_f <- val_f['R2', 'index.corrected']
  q_f <- val_f['Q', 'index.corrected']
  
  val_g <- validate(g2, method = "boot", B = 1000)
  c_g <- val_g['Dxy', 'index.corrected']/2 + 1/2
  r_g <- val_g['R2', 'index.corrected']
   q_g <- val_g['Q', 'index.corrected']
  
   val_h <- validate(h2, method = "boot", B = 1000)
  c_h <- val_h['Dxy', 'index.corrected']/2 + 1/2
  r_h <- val_h['R2', 'index.corrected']
   q_h <- val_h['Q', 'index.corrected']
  
    hr <- summary(g)$coefficients[1,2]
    hr_l <- exp(summary(g)$coefficients[1,1] - summary(g)$coefficients[1,3]*1.96) 
    hr_u <- exp(summary(g)$coefficients[1,1] + summary(g)$coefficients[1,3]*1.96) 
    
     hr_rsi <- summary(h)$coefficients[1,2]
    hr_l_rsi <- exp(summary(h)$coefficients[1,1] - summary(g)$coefficients[1,3]*1.96) 
    hr_u_rsi <- exp(summary(h)$coefficients[1,1] + summary(g)$coefficients[1,3]*1.96) 

    print(typeC)

#colnames(total_df) <- c("HR", "HR_L", "HR_U", "HR_RSI", "HR_L_RSI", "HR_U_RSI", "C_F", "C_G", "C_H","R2_F","R2_G", "R2_H", "Q_F", "Q_G", "Q_H")
  total_df[typeI, 1] = hr
  total_df[typeI, 2] = hr_l
  total_df[typeI, 3] = hr_u
  total_df[typeI, 4] = hr_rsi
  total_df[typeI, 5] = hr_l_rsi
  total_df[typeI, 6] = hr_u_rsi
  total_df[typeI, 7] = c_f
  total_df[typeI, 8] = c_g
  total_df[typeI, 9] = c_h
  total_df[typeI, 10] = r_f
  total_df[typeI, 11] = r_g
  total_df[typeI, 12] = r_h
   total_df[typeI, 13] = q_f
  total_df[typeI, 14] = q_g
  total_df[typeI, 15] = q_h

     d3 <- d2 %>%
        mutate(genetic_score = cut(RSI,
                              breaks = quantile(RSI, probs = c(0, 1/2, 1), na.rm = TRUE),
                              labels = c("Lower Half", "Upper Half"),
                              include.lowest = TRUE))

      d3 <- d3 %>% mutate(genetic_score = as.factor(genetic_score))

      # Create the survival object
      S_PFI <- Surv(d3$PFI.time, d3$PFI)

      # Fit the Kaplan-Meier model using the cut scores
      PFI.kmfit <- survfit(S_PFI ~ genetic_score, data = d3, conf.int = 0.95)

      # Plot the survival curves
      g_plot <- ggsurvplot(fit = PFI.kmfit, # Kaplan-Meier survival model
                 palette = c("pink", "lightblue"), # Custom color palette for three groups
                 legend.title = "Genetic Risk Score", # Legend title
                 legend.labs = c("Lower Half", "Upper Half"), # Label lines
                 title = typeC,
                 xlab = "Time (days)", conf.int = TRUE, pval = TRUE, xlim = c(0,1825), break.time.by = 365
                 )
    
    plots[[typeI]] <- g_plot
  
     typeI = typeI + 1
     
     #pred <- d2 %>% select(case, predictions)
     pred <- d2 %>% select(case, RSI, predictions)
     violin_df_RSI <- left_join(violin_df_RSI, pred, by = join_by(case))
}

#violin_df <- violin_df %>% filter(!is.na(predictions))
violin_df_RSI <- violin_df_RSI %>%
  mutate(RSI = coalesce(RSI.x, RSI.y, RSI.x.x, RSI.y.y, RSI.x.x.x, RSI.y.y.y, RSI.x.x.x.x, RSI.y.y.y.y)) %>%
  dplyr::select(-c(RSI.x, RSI.y, RSI.x.x, RSI.y.y, RSI.x.x.x, RSI.y.y.y, RSI.x.x.x.x, RSI.y.y.y.y))%>%
  mutate(predictions = coalesce(predictions.x, predictions.y, predictions.x.x, predictions.y.y, predictions.x.x.x, predictions.y.y.y, predictions.x.x.x.x, predictions.y.y.y.y)) %>%
  select(-c(predictions.x, predictions.y, predictions.x.x, predictions.y.y, predictions.x.x.x, predictions.y.y.y, predictions.x.x.x.x, predictions.y.y.y.y))
```


###violins
```{r}
violin_df2 <- violin_df_RSI %>% filter(!is.na(RSI))
violin_df2 <- violin_df2 %>% mutate(PFI = as.factor(PFI))

# p <- ggplot(violin_df2, aes(x=interaction(type, PFI), y=predictions, fill = PFI)) + 
#   geom_violin() + xlab("Cancer Type") + ylab("Linear Predictions") + 
#    # stat_summary(fun.data=mean_sdl, mult=1, 
#    #                geom="pointrange", color="black", show.legend = F) +
#   scale_fill_manual(values = c("pink", "lightblue"))
# p


# ggplot(d1, aes(x = PFI, y = predictions, fill = PFI), show.legend = F) +
#   geom_violin(show.legend = F) + xlab("Progression") + ylab("Genetic Prediction") +
#   stat_summary(fun.data=mean_sdl, mult=1,
#                  geom="pointrange", color="black", show.legend = F) +
#   scale_fill_manual(values = c("pink", "lightblue")) +
#   theme_minimal()

p <- ggplot(violin_df2, aes(x=type, y=RSI, fill=PFI)) +
  geom_violin(position = position_dodge(width = 0.8)) +
  geom_boxplot(width = 0.1, position = position_dodge(width = 0.8), outlier.shape = NA, color = "black", show.legend=F) +
  xlab("Cancer Type") +
  ylab("Linear Predictions") +
  labs(title = "RSI by Cancer Type and Outcome") + 
  scale_fill_manual(values = c("pink", "lightblue"))

p

ggsave("violins_RSI.pdf", width = 8, height = 4)


```

###KM
```{r}
myplot <- arrange_ggsurvplots(plots, ncol = 2, nrow = 4) 
#myplot %>% ggsave(device="pdf", filename="survplots.pdf", width = 12, height = 16, units = "in")
#myplot %>% ggsave(device="pdf", filename="survplots_adversarial.pdf", width = 12, height = 16, units = "in")
#myplot %>% ggsave(device="pdf", filename="survplots_logistic.pdf", width = 12, height = 16, units = "in")
myplot %>% ggsave(device="pdf", filename="survplots_RSI.pdf", width = 12, height = 16, units = "in")
```

###plot forest
```{r}
# Extract and reshape manually
forest_data <- total_df[, c("HR", "HR_L", "HR_U", "HR_RSI", "HR_L_RSI", "HR_U_RSI")]
forest_data$term <- rownames(total_df)

# Create two separate data frames, one for each type
orig_df <- forest_data %>%
  select(term, HR, HR_L, HR_U) %>%
  mutate(Type = "Main Analysis") %>%
  rename(HR_point = HR, HR_lower = HR_L, HR_upper = HR_U)

rsi_df <- forest_data %>%
  select(term, HR_RSI, HR_L_RSI, HR_U_RSI) %>%
  mutate(Type = "RSI") %>%
  rename(HR_point = HR_RSI, HR_lower = HR_L_RSI, HR_upper = HR_U_RSI)

# Combine both
long_forest_data <- bind_rows(orig_df, rsi_df)
long_forest_data$Type <- factor(long_forest_data$Type, levels = c("RSI", "Main Analysis"))

# Plot with dodged positions
ggplot(long_forest_data, aes(x = term, y = HR_point, ymin = HR_lower, ymax = HR_upper, color = Type)) +
  geom_point(position = position_dodge(width = 0.5), size = 2) +
  geom_errorbar(position = position_dodge(width = 0.5), width = 0.2) +
  geom_hline(yintercept = 1, color = "black") +
  scale_x_discrete(limits = rev(rownames(total_df))) +
  scale_color_manual(
    values = c("Main Analysis" = "blue", "RSI" = "red"),   # choose your colors
    breaks = c("Main Analysis", "RSI"),                    # desired legend order
    name = "Model"
  )+
  coord_flip() +
  xlab(" ") +
  ylab("Hazard Ratio (95% Confidence Interval)") +
  labs(title = "Per-Cancer Prognostic Score Hazard Ratios", color = "Model") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    axis.text.x = element_text(size = 12),
    axis.text.y = element_text(size = 12),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 13)) 

ggsave("forest_RSI.pdf", width = 8, height = 5)
```
###CRQ
```{r}
library(gridExtra)
c_df <- total_df %>% dplyr::select(c("C_F", "C_G", "C_H"))
c_df <- c_df %>% mutate(types = rownames(c_df))
df_long <- c_df %>%
  pivot_longer(cols = c(C_F, C_G, C_H), names_to = "Group", values_to = "Value") 
# Create the grouped bar plot
c_plot <- ggplot(df_long, aes(x = types, y = Value, fill = Group)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +  # Grouped bars
  scale_fill_manual(values = c("C_F" = "lightblue", "C_G" = "pink", "C_H" = "lightgreen"), labels = c("Clinical", "Clinical + Genomic", "Clinical + RSI")) +  # Custom colors
  labs(title = "C Statistic With and Without Prognostic Scores",
       x = "Cancer Type",
       y = "Corrected C Statistic",
       fill = "Model") +
 
  theme_minimal()+ 
  coord_cartesian(ylim = c(0.5,0.7))


c_df <- total_df %>% dplyr::select(c("R2_F", "R2_G", "R2_H"))
c_df <- c_df %>% mutate(types = rownames(c_df))
df_long <- c_df %>%
  pivot_longer(cols = c(R2_F, R2_G, R2_H), names_to = "Group", values_to = "Value") 
# Create the grouped bar plot
r_plot <- ggplot(df_long, aes(x = types, y = Value, fill = Group)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +  # Grouped bars
  scale_fill_manual(values = c("R2_F" = "lightblue", "R2_G" = "pink", "R2_H" = "lightgreen"), labels = c("Clinical", "Clinical + Genomic", "Clinical + RSI")) +  # Custom colors
  labs(title = expression(paste("R"^2," Values With and Without Prognostic Scores")),
       x = "Cancer Type",
       y = expression("Corrected R"^2),
       fill = "Model") +
  theme_minimal()+ 
  coord_cartesian(ylim = c(-0.04,0.15)) #normal


c_df <- total_df %>% dplyr::select(c("Q_F", "Q_G", "Q_H"))
c_df <- c_df %>% mutate(types = rownames(c_df))
df_long <- c_df %>%
  pivot_longer(cols = c(Q_F, Q_G, Q_H), names_to = "Group", values_to = "Value") 
# Create the grouped bar plot
q_plot <- ggplot(df_long, aes(x = types, y = Value, fill = Group)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +  # Grouped bars
  scale_fill_manual(values = c("Q_F" = "lightblue", "Q_G" = "pink", "Q_H" = "lightgreen"), labels = c("Clinical", "Clinical + Genomic", "Clinical + RSI")) +  # Custom colors
  labs(title = "Q Values With and Without Prognostic Scores",
       x = "Cancer Type",
       y = "Corrected C Statistic",
       fill = "Model") +
 
  theme_minimal()+ 
  coord_cartesian(ylim = c(-0.015,0.03)) #normal


g <- grid.arrange(c_plot, r_plot, q_plot, nrow = 3)

#ggsave("C_R_Q_plot.pdf", g, width = 8, height = 11 )
#ggsave("C_R_Q_plot_adversarial.pdf", g, width = 8, height = 11 )
#ggsave("C_R_Q_plot_logistic.pdf", g, width = 8, height = 11 )
ggsave("C_R_Q_plot_RSI.pdf", g, width = 8, height = 11 )
```